---
title: "vri data cleaning"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
##############
READ ME
This script pulls out the relevant vri attributes for this project from the raw
object that contains 193 attributes and does some formatting. 

FUTURE IMPROVEMENTS
- Could download latest vri release (2023)
- Could look at other VRI layers for height info and for BA height estimates
##############

Load packages and read in data
```{r}
rm(list = ls(all=TRUE))

library(tidyverse)
library(here)

#Read in raw vri data
vri <- read_csv(here("./data/raw/vri_hdm_sites.csv"))
dim(vri)

#Read in site data
site <- read_csv(here("./data/cleaned/site_data.csv"))
```

Pull out relevant variables. We want: a vri id column, site_id (of associated
hdm site), projected date, species, their relative proportions, age, and height.
```{r}
vri <- vri %>% 
  select(FEATURE_ID,
         site_id,
         tree_type,
         PROJECTED_DATE,
         starts_with("BEC"),
         HARVEST_DATE,
         starts_with("SPECIES_CD"),
         starts_with("SPECIES_PCT"),
         starts_with("PROJ_AGE"),
         starts_with("PROJ_HEIGHT"))
```

Reformat dates
```{r}
#Convert date columns in vri data to date classes. Raw format is month-day-year
#hour:min:sec
vri <- vri %>% 
  mutate(PROJECTED_DATE = mdy_hms(PROJECTED_DATE),
         HARVEST_DATE = mdy_hms(HARVEST_DATE))

#Pull out the year of these
vri <- vri %>% 
  mutate(PROJECTED_YR = year(PROJECTED_DATE),
         HARVEST_YR = year(HARVEST_DATE)) %>% 
  relocate(PROJECTED_YR, .after = PROJECTED_DATE) %>% 
  relocate(HARVEST_YR, .after = HARVEST_DATE)

#Convert date columns in site data to date classes
site <- site %>% rename(date_sur_char = date_surveyed) %>% 
  mutate(date_sur = dmy(date_sur_char)) %>% 
  mutate(yr_sur = year(date_sur)) %>% 
  relocate(c(date_sur, yr_sur), .after = date_sur_char)

#Add the year surveyed to vri data
x <- site %>% 
  select(site_id, yr_sur)
vri <- left_join(vri, x, by = "site_id")

#Calculate the projected age at the time of measurement:
vri <- vri %>% 
  mutate(YR_DIFF = yr_sur - PROJECTED_YR) %>% 
  mutate(AGE_MEAS_1 = PROJ_AGE_1 + YR_DIFF,
         AGE_MEAS_2 = PROJ_AGE_2 + YR_DIFF)

#Pull out estimates of Hw relative abundance, age and height
#HW is always first or second species, which is why the logic in this code 
#works
#Note: age we are easily able to correct to the height at the measurement 
#year but height we can't. It is projected to 2021.
vri <- vri %>% 
  mutate(PERC_HW = if_else(SPECIES_CD_1 == "HW", SPECIES_PCT_1, 
                           SPECIES_PCT_2), 
         AGE_HW = if_else(SPECIES_CD_1 == "HW", AGE_MEAS_1, 
                          AGE_MEAS_2),
         HEIGHT_HW = if_else(SPECIES_CD_1 == "HW", PROJ_HEIGHT_1, 
                             PROJ_HEIGHT_2))

#Has everything we need to start modelling. Export this: 
write_csv(vri, here("./data/cleaned/vri_c.csv"))
```

