---
title: "02_define_dmr"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
This script defines two, tree-level versions of dwarf misltetoe rating (DMR) 
from the crown-third level DMR ratings and other dwarf misltetoe tree data.

The output (./data/workflow/trees_mapped.csv) of the script that adds 
georeferenced locations to each tree (./scripts/01_convert_dist_az_to_point.R) 
is the input for this script. The output from this script feeds into the script 
that simulates trees to create a dataset with complete data for a single 
standard transect length (./scripts/03_simulate_trees.Rmd). 

Load packages
```{r}
library(tidyverse)
library(here)
```

Read in data.
```{r}
rm(list=ls(all=TRUE))

trees <- read_csv(here("./data/workflow/trees_mapped.csv"))
```

#Traditional DMR
Traditionally, tree-level DMR is calculated as the sum of each crown-third DMR
rating. Crown third ratings record the proportion of primary branches infected
within each third of the live crown: 0 (none), 1 (>0% and ≤50%), 2 (>50%).

Calculate that. 
```{r}
trees <- trees %>% 
  mutate(dmr=dmr_l+dmr_m+dmr_u)
trees %>% select(dmr) %>% summary()
```


#Adapted DMR
An adapted version of DMR was created to meet two needs:

(1) Traditional DMR ignores infections below the live crown of live trees. 
These infections are typically inactive, but I decided it was informative to 
distinguish trees with an infection below the live crown (but not in the live
crown) from trees that have no infections. 
(2) A variable that describes HDM infection in live and dead hemlock trees is
useful for creating plots. Traditional DMR only includes live trees (and the 
modelling anlyses for this project are also restricted to live trees), but dead
trees are given distinct classes in this variable so it can double as a 
plotting variable. 

Variable logic: 
- Tree not a hemlock or amabilis fir: "-"
- Tree a hemlock or amabilis fir AND dead AND infected: "DI"
- Tree a hemlock or amabilis fir AND dead AND uninfected: "DU"
- Tree a hemlock or amabilis fir AND alive AND infected below live crown but
no infections in live crown: "IBLC"
- Tree a hemlock or amabilis fir AND alive AND with infection in live crown: 
traditional DMR (calculated above)

Calculate it.
```{r}
#Calculate new factor version of tree level dmr
trees <- trees %>% 
  mutate(dmr_f = case_when(!spp %in% c("Hw", "Ba") ~ "-",
                           (spp %in% c("Hw", "Ba") & 
                              status %in% c("DS", "SN") & hdm_pa=="Y") ~ "DI", 
                           (spp %in% c("Hw", "Ba") & 
                              status %in% c("DS", "SN") & 
                              hdm_pa %in% c("N", "U")) ~ "DU",
                           (spp %in% c("Hw", "Ba") & 
                              hdm_pa=="Y" & b_lc=="Y" & dmr==0) ~ "IBLC",
                           TRUE ~ as.character(dmr))) %>% 
  mutate(dmr_f = factor(dmr_f, levels = c("-", "DU", "0", "DI", "IBLC", 
                                          "1", "2", "3", "4", "5", "6")))

trees %>% select(dmr_f) %>% table(useNA = "ifany") #Looks good!
```

#Export
```{r}
#Format so tree-level DMR variables are next to crown third DMR variables
trees <- trees %>% 
  relocate(c(dmr, dmr_f), .before = broom_pa)

#Write it to a csv
write_csv(trees, here("./data/workflow/trees_dmr.csv"))
```
