---
title: "simulate trees"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
##############
READ ME
This script simulates data to extend the regen transects to a standard length
of 50m. This came out of the choice to use variable length transects. Transect 
length was determined by the infected Hw farthest from the mature edge at a 
site; if that tree was <15m from the edge, we measured to 15m as a standard 
minimum. The longest transect length in the dataset is 51m (cr_3) and the 
shortest is 15m (the minimum). 

All trees beyond the transect end should be uninfected. This was measured in a
sense - individual trees weren't measured but the transect end corresponds to 
the infection perimeter of HDM. We are using this assumption to extend the 
data.

The advantage of extending the data is that it allows us to make use of the 
data from sites that were measured >15m from the edge. If we used data from 
those sites without simulating data for the other sites, it would bias the 
predictive models towards overpredicting HDM infection. 

The simulations are simple. The number of stems/ha of each spp - status -
crown class combination is determined on each transect. These transect level 
counts are averaged across the three transects at a site, then used to generate
a corresponding number of points to fill the area on each transect between its 
end and 50m.

This script starts with the ouput object from the stem map script (stem map.R).
The output feeds into the script to estimate crown volume (crown volume.Rmd)
##############

```{r}
#Read in packages
library(tidyverse)
library(here)
library(sf)
library(tmap)
```


#Read in data
```{r}
rm(list=ls(all=TRUE))

####Dataset 1: hdm site data 
#This contains variables at the site level
site <- read_csv(here("./data/cleaned/site data.csv"))
str(site)
summary(site)

#Create a single variable for BEC
site <- site %>% unite(col = bec, sep = "", 
                       c("bec_z", "bec_sz", "bec_var"), 
                       remove = FALSE) %>% 
  mutate(bec = factor(bec))

#Convert date_surveyed to date format
site <- site %>% rename(date_sur_char = date_surveyed) %>% 
  mutate(date_sur = dmy(date_sur_char)) %>% 
  relocate(date_sur, .after = date_sur_char)
#Extract year
site <- site %>% mutate(yr_sur = year(date_sur))
#Recalculate age column
site <- site %>% mutate(age = yr_sur - yr_har)

####Dataset 2: transect data
transect <- read_csv('./data/cleaned/transect data_c.csv')
str(transect)

####Dataset 3: main tree dataset
#Read in the tree data
trees <- read_csv(here("data/workflow/trees_mapped.csv"))

#Convert factor vars to factors: 
trees <- trees %>% mutate(across(
  c(site_id, spp, status, hdm_pa, b_lc, 
    broom_pa, broom_pos, stem_pa, crown_class, crown_cond, outside_10, 
    assessed_by, tree_type, dmr_f),
  ~as.factor(.)))

#Convert plot_id and dmrs to integers
trees <- trees %>% mutate(across(
  c(plot_id, dmr_l, dmr_m, dmr_u, dmr), ~as.integer(.)))

#Check variables in object
str(trees)
```

#Figure out what we need to simulate
Need to know two things: 
(1) what is transect area that is unmeasured at each site?
(2) how many trees are there on average

Starting with the first piece
```{r}
#Goal is to simulate data from where a transect ends to a distance of 50m. 
#Get a summary of the transect length
summary(transect$tr_leng)

#Calculate the measured transect area
transect <- transect %>% 
  mutate(tr_area_m2 = tr_leng*5,
         tr_area_ha = tr_area_m2/10000)

#Calculate how length is missing for each transect. 
transect <- transect %>% 
  mutate(mis_dist = 50 - tr_leng)

#Convert that to area (m2) and then hectares
transect <- transect %>% 
  mutate(mis_area_m2 = mis_dist*5,
         mis_area_ha = mis_area_m2/10000)
summary(transect$mis_area_ha)

#Account for cr_3 site that was measured to 51m. If missing area is < 0, set
#it to 0
transect <- transect %>% 
  mutate(across(starts_with("mis"), ~ if_else(. < 0, 0, .)))
transect %>% select(starts_with("mis")) %>% 
  summary()
```

Now calculate the second piece. Group at the transect level by species, status 
and crown class. Then average across the three transects to get numbers to 
apply to each missing transect area within a site.
```{r}
#Filter to regen trees
regen <- trees %>% 
  filter(tree_type == "regen")

#Calculate a per hectare factor at the tree level
#Join the transect area
x <- transect %>% select(transect_id, tr_area_ha)
regen <- left_join(regen, x, by = join_by(plot_id == transect_id))

#Calculate per hectare factor = 1/transect area
regen <- regen %>% 
  mutate(phf_tr = 1/tr_area_ha)

#Filter out regen that are immediately next to the edge (within 5m)
regen <- regen %>% 
  filter(dist_y_h > 5)

#Redefine species factor. If species were only in mature, we don't need those
#factor levels anymore
regen <- regen %>% 
  mutate(spp = as.character(spp)) %>% 
  mutate(spp = as.factor(spp))
levels(regen$spp)

#Redefine plot_id as a factor, this allows us to force functions to create rows
#for situations for when there aren't any cases of it in the data
regen <- regen %>% 
  mutate(plot_id = as.factor(plot_id))

#Now get summary at the transect level
#How many combinations should there be? 
length(unique(regen$plot_id)) #33 transects
length(levels(regen$spp)) #10 spp
length(levels(regen$status)) #6 status levels
length(levels(regen$crown_class)) # 5 crown class levels

#Crown class is only defined for live trees. There were a few cases where this
#was recorded for recently dead trees, because I thought it might be useful. 
#But we can ignore these and reassign them "-" for crown class. For live trees, #all trees should have a crown class value, so cases where crown_class is "-" 
#can be ignored. 

#This means live and dead trees should be summarised separately, then 
#recombined. The number of combinations should be: 
##Live trees: 4 crown class levels x 3 status levels x 10 spp x 33 transects 
## == 3960 cases
##Dead trees: 3 status levels x 10 spp x 33 transects == 990
##Total: 3960 live tree cases + 990 dead tree cases == 4950

#Separate live and dead trees
live <- regen %>% 
  filter(status %in% c("LS", "LL", "LF"))
dead <- regen %>% 
  filter(status %in% c("RD", "DS", "SN"))

#Redefine status and crown class in both datasets. 
#This should remove "-" crown class level for livetrees and set "-" as the 
#only level for dead trees. 
live <- live %>% 
  mutate(across(c(status, crown_class), ~ as.character(.x))) %>% 
  mutate(across(c(status, crown_class), ~ as.factor(.x))) 
levels(live$status)
levels(live$crown_class)

dead <- dead %>% 
  mutate(across(c(status, crown_class), ~ as.character(.x))) %>% 
  mutate(crown_class = "-") %>% 
  mutate(across(c(status, crown_class), ~ as.factor(.x)))
levels(dead$status)
levels(dead$crown_class)

#Calculate the summary
#Fill cases that are missing with 0 for the stem counts and NA for dbh metrics
#(if there are no trees, there shouldn't be a value for mean or sd of dbh)
live_comp <- live%>% 
  group_by(plot_id, spp, status, crown_class)  %>% 
  summarise(n_stem_tr = n(),
            n_stem_ha = sum(phf_tr),
            dbh_mean = mean(dbh, na.rm = T),
            dbh_sd = sd(dbh, na.rm = T)) %>% 
  ungroup() %>% 
  complete(plot_id, spp, status, crown_class, 
           fill = list(n_stem_tr = 0,
                       n_stem_ha = 0))
dim(live_comp)

dead_comp <- dead %>% 
  group_by(plot_id, spp, status, crown_class)  %>% 
  summarise(n_stem_tr = n(),
            n_stem_ha = sum(phf_tr),
            dbh_mean = mean(dbh, na.rm = T),
            dbh_sd = sd(dbh, na.rm = T)) %>% 
  ungroup() %>% 
  complete(plot_id, spp, status, crown_class, 
           fill = list(n_stem_tr = 0,
                       n_stem_ha = 0))
dim(dead_comp)

#Add site id
x <- transect %>% select(site_id, transect_id)
live_comp <- live_comp %>% 
  mutate(plot_id = as.character(plot_id)) %>% 
  mutate(plot_id = as.integer(plot_id))
live_comp <- left_join(live_comp, x, by = join_by(plot_id == transect_id))

dead_comp <- dead_comp %>% 
  mutate(plot_id = as.character(plot_id)) %>% 
  mutate(plot_id = as.integer(plot_id))
dead_comp <- left_join(dead_comp, x, by = join_by(plot_id == transect_id))

#Recombine the live and dead datasets
tree_comp <- rbind(live_comp, dead_comp)

#Check levels of factors
levels(tree_comp$spp)
levels(tree_comp$status)
levels(tree_comp$crown_class)

#Summarise to the site level, averaging across the three transects
#Should be 1650 cases 
tree_comp <- tree_comp %>% 
  group_by(site_id, spp, status, crown_class) %>% 
  summarise(n_stem_tr = mean(n_stem_tr, na.rm = T),
            n_stem_ha = mean(n_stem_ha, na.rm = T),
            dbh_mean = mean(dbh_mean, na.rm = T),
            dbh_sd = mean(dbh_sd, na.rm = T))
dim(tree_comp)

#Look at hemlock case to make sure this all worked
#Looks good. 
x <- tree_comp %>% filter(spp == "Hw")
```

Now combine the two pieces to get a round number of the number of trees of each
species, status and crown class combination that we need to generate for each
transect. 
```{r}
#Create new dataset structured at the transect level. For each transect it
#should have the area that needs to be simulated and rows for spp, status, 
#crown class combination created above. 
x <- transect %>% select(site_id, transect_id, tr_leng, mis_area_ha)
sim_needs <- full_join(x, tree_comp, by = "site_id", 
                       relationship = "many-to-many")

#Calculate the number of trees that need to simulated
sim_needs <- sim_needs %>% 
  mutate(n_stem_sim = round(n_stem_ha*mis_area_ha, digits = 0))
```

#Generate new trees
The output of the last section is a dataframe that has all combinations of spp,
status and crown class for all transects; for each of these rows, it has a 
column that indicates how many trees need to be simulated in each case. The next step is to expand that count to an actual dataframe of simulated trees. 

Write a function that takes each row in the dataframe, and creates a the 
corresponding number of trees. 
```{r}
#Input data: dataframe described above
#Function operates rowwise, analgous to a loop
#For each row it creates a new variable - a list with a single item, a 
#dataframe of simulated trees. 
#That dataframe will have n_stem_sim rows
##site_id, transect_id, spp, status and crown class are replicated
##dbh is sampled from a normal distribution if there was >1 trees in that
##spp-status-crown class combination at that site
##dist_y_h is sampled from the end of the transect to 50m, with replacement
##dist_x is sampled from -2.5 to 2.5
##with replacement allows trees to be in the same place

f_sim_trees <- function(df) {
  set.seed(11)
  
  df %>%
    rowwise() %>%
    mutate(sim_trees = list(
      tibble(
        dbh = if (!is.na(dbh_sd)) {
        rnorm(n_stem_sim, dbh_mean, dbh_sd)
      } else {
        rep(dbh_mean, n_stem_sim)
      },
        dist_x = sample(-2.5:2.5, n_stem_sim, replace = TRUE),
        dist_y_h = sample(tr_leng:50, n_stem_sim, replace = TRUE)))) %>%
    unnest(sim_trees)  # Unnest the simulated data into rows
}
```

Now simulate the trees
```{r}
#Filter to cases where we need to simulate trees
sim_needs <- sim_needs %>% filter(n_stem_sim>0)
dim(sim_needs) #293 cases

#How many trees do we need to simulate total? 
sum(sim_needs$n_stem_sim) #954

#Run the function
sim_trees <- f_sim_trees(sim_needs)

#Plot to makes sure this is doing what we want: 
ggplot(sim_trees, aes(x = crown_class, y = dbh, colour = site_id)) +
  geom_point()
ggplot(sim_trees, aes(x = crown_class, y = dist_y_h)) + 
  geom_point()
```

#Align simulate points with the larger dataset
Need to do two things: 
(1) add columns to align it with the rest of the data
(2) map the points from the transect starts

```{r}
#First see which columns are in the measured data that aren't in the simulated
#data
setdiff(names(trees), names(sim_trees))

#Variables that need  values: 
##X, Y, plot_x_utm, plot_y_utm, will get value in funtion that maps these to coordinates
##tree_id, assign as ascending from s1 to sx (x=total num of simulated trees)
##plot_id, rename transect_id plot_id
##all hdm variables, given values corresponding to no infection
##tree_type, all regen

#Variables that can be left NA:
##flag_id
##dist_y, slope distance, we don't need it
##height_m
##dam_agent
##path_ind
##crown_cond
##assessed_by
##ht_corr
##Dec
##corr_az_deg
##tr_az
##tr_area_ha
##phf_tr

#Create tree id
sim_trees <- sim_trees %>% 
  mutate(tree_id = str_c("s", row_number(), sep = "")) %>% 
  relocate(tree_id, .before = site_id)

#Rename transect_id
sim_trees <- sim_trees %>% 
  rename(plot_id = transect_id)

#Assign tree_type
sim_trees <- sim_trees %>% 
  mutate(tree_type = "regen")

#Assign hdm variables
#Rules: 
##If spp == Hw or Ba and status == LS, LL, LF or RD, hdm_pa = N, other factors 
##are "-", crown third dmrs are 0 and tree level dmr is 0
##If spp == Hw or Ba and status == DS or SN, hdm_pa = N, other factors 
##are "-" and dmrs are NA
##If spp != Hw or Ba, hdm_pa = "-", all other factors are "-" and dmrs are NA
x1 <- sim_trees %>% 
  filter(spp %in% c("Hw", "Ba") &
           status %in% c("LS", "LL", "LF", "RD")) %>% 
  mutate(hdm_pa = "N",
         b_lc = "-",
         dmr_l = 0,
         dmr_m = 0,
         dmr_u = 0,
         broom_pa = "-",
         broom_pos = "-",
         stem_pa = "-",
         dmr = 0,
         dmr_f = "0")

x2 <- sim_trees %>% 
  filter(spp %in% c("Hw", "Ba") &
           status %in% c("DS", "SN")) %>% 
  mutate(hdm_pa = "N",
         b_lc = "-",
         dmr_l = NA,
         dmr_m = NA,
         dmr_u = NA,
         broom_pa = "-",
         broom_pos = "-",
         stem_pa = "-",
         dmr = NA,
         dmr_f = "DU")

x3 <- sim_trees %>% 
  filter(!(spp %in% c("Hw", "Ba"))) %>% 
  mutate(hdm_pa = "-",
         b_lc = "-",
         dmr_l = NA,
         dmr_m = NA,
         dmr_u = NA,
         broom_pa = "-",
         broom_pos = "-",
         stem_pa = "-",
         dmr = NA,
         dmr_f = "-")
sim_trees <- rbind(x1, x2, x3)

#Remove variables created for simulating the trees
str(sim_trees)
sim_trees <- sim_trees %>% 
  select(-c(mis_area_ha, n_stem_tr, n_stem_ha, n_stem_sim, dbh_mean,
            dbh_sd))
str(sim_trees)
```

Define the function to map trees from transect start points
```{r}
#azimuth = azimuth from centre point in degrees
#distance = distance from center point in m
#xcenter/ycenter = X and Y coordinates of trimble reference point
#crs = EPSG code of Coordinate reference system you are using
polar_to_XY <- function(azimuth,
                        distance,
                        xcenter,
                        ycenter,
                        shape_file = TRUE,
                        crs) {
  
  if((max(azimuth) - min(azimuth) < 2*pi) == TRUE){
    print("WARNING: This function assumes azimuth is in degrees, please check")
  }
  angle = azimuth * pi/180
  #Convert to radians
  #angle = angle*pi/180
  angle = 2*pi - (angle - pi/2)
  x = xcenter + distance * cos(angle)
  y = ycenter + distance * sin(angle)
  
  #define output point locations
  tree_locations <- data.frame(X = x, Y = y)
  
  if(shape_file == T){# output a shape file of the tree locations
    print(paste("creating spatial points layer from tree locations. CRS is:", crs))
    
    tree_locations <- sf::st_as_sf(tree_locations, coords = c("X", 'Y'), crs = crs)
    
  }
  else{
    tree_locations
  }
  return(tree_locations)
}
```

Do the mapping
```{r}
#Read in the trimble points that have the transect start points in them
trimb <- read_csv(here('./data/workflow/trimb_radjusted.csv'))

#Subset to just the points used for stem mapping and then to transect start points
trimb <- trimb %>% 
  filter(pt_type == "tran start" & stem_map == "Y")

#Calculate distance and azimuth of each simulated tree from the transect start
##This is a right triangle with sides: dist_x, dist_y_h.

##Add transect azimuth to the sim_trees dataset
x <- transect %>% select(transect_id, tr_az)
sim_trees <- left_join(sim_trees, x, by = join_by(plot_id == transect_id))

##Calculate distance
sim_trees <- sim_trees %>% 
  mutate(dist_m = (dist_x^2 + dist_y_h^2)^(0.5))
summary(sim_trees$dist_m)

##Calculate angle adjustment relative to transect azimuth:
sim_trees <- sim_trees %>% 
  mutate(az_adj = atan(dist_x/dist_y_h)*(180/pi))

##Calculate actual azimuth (0-360 deg)
sim_trees <- sim_trees %>% 
  mutate(corr_az_deg = (tr_az + az_adj)%%360)

#Add the transect start coordinates from the trimb object
x <- trimb %>% 
  select(pt_id, plot_x_utm, plot_y_utm)
sim_trees <- left_join(sim_trees, x, by = join_by(plot_id == pt_id))

#Set the cordinate reference system
crs <- 3005

# Run the function. Need to set azimuth, distance, xcenter, ycenter equal to 
#variable names
sim_trees <- sim_trees %>%
  mutate(polar_to_XY(azimuth = corr_az_deg, distance = dist_m,
                     xcenter = plot_x_utm, ycenter = plot_y_utm, crs = crs,
                     shape_file = TRUE))

class(sim_trees)

#Convert it to a spatial object:
sim_trees <- st_as_sf(sim_trees)

#Plot to make sure it worked as expected: 
tmap_mode("plot") +
  tm_shape(sim_trees, is.master = TRUE) + 
    tm_symbols(col = "tree_type") +
  tm_facets("site_id")

#Extract the coordinates as columns called X and Y
sim_trees <- sim_trees %>%
  mutate(X = st_coordinates(.)[,1],
         Y = st_coordinates(.)[,2])  

# Convert to a dataframe (drop the geometry)
sim_trees <- st_drop_geometry(sim_trees)
```

Do final alignments, then join the simulated trees to the larger tree dataset
```{r}
#First see which columns are in the measured data that aren't in the simulated
#data
#Looks good. All columns that can be NA. 
setdiff(names(trees), names(sim_trees))

#Look at which columns are in the simulated data that aren't in the measured
#data. 
#Just one column created in data processing. Drop it. 
setdiff(names(sim_trees), names(trees))
sim_trees <- sim_trees %>% 
  select(-az_adj)

#Add columns to sim_trees and set NA so it aligns with trees
missing_cols <- setdiff(names(trees), names(sim_trees))
for (col in missing_cols) {
  sim_trees[[col]] <- NA
}
#Check. All columns the same. Good
setdiff(names(trees), names(sim_trees))
setdiff(names(sim_trees), names(trees))

#Add a column that identies simulated trees
sim_trees <- sim_trees %>% 
  mutate(sim_tree = "Y")
trees <- trees %>% 
  mutate(sim_tree = "N")

#Bind the two dataframes
trees <- rbind(trees, sim_trees)
```

#Plot the stem maps
A final step to make sure this all makes sense
```{r}
#Make another tree type varaible for plotting that combines tree type and 
#whether a tree was simulated. 
trees <- trees %>% 
  mutate(tree_type_2 = case_when(
    tree_type == "regen" & sim_tree == "Y" ~ "sim regen",
    tree_type == "regen" & sim_tree == "N" ~ "meas regen",
    tree_type == "mature" ~ "mature"))

#Convert trees object to sf object
trees_sf <- st_as_sf(trees, coords = c("X", "Y"), crs = 3005)

#Plot points, coloured by tree_type_2
tmap_mode("plot") +
  tm_shape(trees_sf, is.master = TRUE) + 
    tm_symbols(col="tree_type_2") + 
  tm_facets("site_id")

#Plot points, coloured by tree_type_2, sized by dbh
tmap_mode("plot") +
  tm_shape(trees_sf, is.master = TRUE) + 
    tm_symbols(col="tree_type_2", size = "dbh", scale = 2) + 
  tm_facets("site_id")

#Plot points, coloured by tree_type_2, sized by dbh, shape = crown_class
tmap_mode("plot") +
  tm_shape(trees_sf, is.master = TRUE) + 
    tm_symbols(col="tree_type_2", shape = "crown_class", size = "dbh", 
               scale = 2) + 
  tm_facets("site_id")

#Plot points, coloured by spp
tmap_mode("plot") +
  tm_shape(trees_sf, is.master = TRUE) + 
    tm_symbols(col= "spp", size = "dbh", scale = 2) + 
  tm_facets("site_id")

#Check Ba is only at one site
tmap_mode("plot") +
  tm_shape(filter(trees_sf %>% filter(spp == "Ba")), is.master = TRUE) + 
    tm_symbols(col= "spp", size = "dbh", scale = 2) + 
  tm_facets("site_id")

#Coloured by DMR, shape = tree_type_2
tmap_mode("plot") +
  tm_shape(trees_sf, is.master = TRUE) + 
    tm_symbols(col= "dmr_f", size = "dbh", scale = 2, shape = "tree_type_2") + 
  tm_facets("site_id")
```

#Export
Everything looks good. Export object as a csv. 
```{r}
# write_csv(trees, here("./data/workflow/trees_sim.csv"))
```

