---
title: "Site Level Analyses"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
WHAT?
Script for running site level analyses in the general form (spread measure ~ site variable).

UPDATED: 28 Jun 2024

#LOAD PACKAGES
```{r}
rm(list=ls(all=TRUE))

#Load packages
library(tidyverse)
library(RColorBrewer)
library(ggrepel)
library(patchwork)
```

#DATA PREP

Read in and format the site and tree datasets
```{r}
####Read in data and inspect it
##Dataset 1: site data. 
#This contains variables at the site level. The key pieces for analysis are: year harvested and CWH subzone
site <- read_csv("data/cleaned/site data.csv")
str(site)
summary(site)

##Create some variables to use in summary tables and figures
#Calculate approximate age of regen. Sites assessed in 2023 so age = 2023 - year harvested
site <- site %>% mutate(age = 2023-yr_har)
site$age #luckily, got pretty good spread

#Create a single variable from all the pieces of a BEC zone
site <- site %>% unite(col = bec, sep = "", c("bec_z", "bec_sz", "bec_var"), remove = FALSE)

#Create a variable that combines site_id, bec, and regen age 
site <- site %>% unite(col = id_bec_age_f, sep = ", ", 
                       c("site_id", "bec", "age"), 
                       remove = FALSE) %>% 
  mutate(id_bec_age_f = factor(id_bec_age_f))

#Make cluster, bec and id_bec_age_f factors
site <- site %>% mutate(across(c(cluster, bec, id_bec_age_f), ~as.factor(.)))
str(site)

##Dataset 2: tree data
#This contains data for each measured tree (mature and regen trees). 
#Treats assessed_by var as a logical for some reason and throws up an error. Not a problem and not going to use this var
trees <- read_csv("data/workflow/trees_mapped.csv")
str(trees)
summary(trees)

#Convert factor vars to factors: 
trees <- trees %>% mutate(across(
  c(site_id, spp, status, hdm_pa, b_lc, 
    broom_pa, broom_pos, stem_pa, crown_class, crown_cond, outside_10, assessed_by, tree_type, dmr_f),
  ~as.factor(.)))

#Convert plot_id and dmrs to integers
trees <- trees %>% mutate(across(
  c(plot_id, dmr_l, dmr_m, dmr_u, dmr), ~as.integer(.)))

str(trees)

#Create new factor dmr variable that has fewer levels and ascends logically from healthy to most infected
levels(trees$dmr_f)
trees <- trees %>% mutate(dmr_f2 = case_when((dmr_f =="1" | dmr_f=="2") ~ "1-2", (dmr_f =="3" | dmr_f=="4") ~ "3-4", (dmr_f =="5" | 
                                    dmr_f=="6") ~ "5-6", .default = dmr_f))
trees <- trees %>% mutate(dmr_f2 = factor(dmr_f2, levels = c("-", "DU", "0","IBLC", "1-2", "3-4", "5-6", "DI")))
summary(trees$dmr_f2)

#Join some site level variables to tree data: 
trees <- left_join(trees, select(site, cluster, 
                                 bec, id_bec_age_f, age, site_id), by="site_id")
```

Define some consistent colours to plot with
```{r}
#spp colours Ba, Cw, Fd, Hw
colors_spp = c("burlywood1", "firebrick", "burlywood4", "dodgerblue")

#hdm_pa colours. Didn't include a colour for non-Hw trees because not creating any plots with them and this variable here. 
levels(trees$hdm_pa)
colors_hdm_pa <- c("dodgerblue", "grey", "goldenrod1")

#dmr_f2 colours. Didn't include a colour for non-Hw trees because not creating any plots with them and this variable here. 
levels(trees$dmr_f2)
colors_dmr_f2 = c("dodgerblue4", "dodgerblue", "khaki1", "khaki3", "goldenrod1", "goldenrod3", "hotpink")
```


#SECTION 1: Site climate comparisons
Goal of this section is to descriptively describe differences in climate variables between sites that might influence HDM biology.
```{r}
####Read in climate data
#Data downloaded from ClimateBC (https://climatebc.ca/mapVersion, 24 Jun 2024). ClimateBCv7.50. Period: Normal_1991_2020
climate <- read_csv('./data/raw/hdm_climdata.csv') #raw data from climateBC

####Check coordinates were inputted correctly, then join site_ids 
#Sites are identified by coordinates in the climate dataset. Check they are right before joining site_ids
site <- site %>% mutate(dd_X_r = round(dd_X, digits = 4), #round number of decimals to match climateBC rounding
                        dd_Y_r = round(dd_Y, digits = 5))
x <- site$dd_X_r - climate$long
y <- site$dd_Y_r - climate$Lat
print(x); print(y)
#Good, almost al 0s, just one that is different by 1 x 10e-4 (a rounding error)

#Join site_id, bec and cluster by latitude values
climate <- left_join(climate, select(site, site_id, cluster, bec, dd_Y_r), by = c("Lat" = "dd_Y_r"))

#move site_id to the front
climate <- climate %>% select(site_id, cluster, bec, everything())

#### Define climate variables that were in Bianca's paper
#Descriptions of these variables and whether they are exact matches or approximiations of variables used in paper are in: /Users/hannosoutham/Library/CloudStorage/OneDrive-UBC(1)/Msc/Thesis/ch1.docx

#1 Growing degree days
climate <- climate %>% rowwise() %>% mutate(DD5_gs = sum(c_across(DD5_sp:DD5_at)))
select(climate, DD5, DD5_gs)

#2 Growing season radians
climate <- climate %>% rowwise() %>% mutate(Rad_gs = mean(c_across(Rad_sp:Rad_at)))
select(climate, MAR, Rad_gs)

#3 Lowest average minimum monthly temperature (Tmin_wt)
# find lowest Tmin value across all months
climate <- climate %>% rowwise() %>% mutate(Tmin_all = min(c_across(Tmin_01:Tmin_12)))
select(climate, Tmin_wt, Tmin_all)

#4 Standard deviation of lowest average minimum monthly temperature  
# calculate the standard deviation in Tmin across all of the months
climate <- climate %>% rowwise() %>% mutate(Tmin_sd = sd(c_across(Tmin_01:Tmin_12)))
select(climate, Tmin_sd)

#5 Snow (PAS) 
select(climate, MAP, PAS)

#6 Rain
climate <- climate %>% mutate(PAR = MAP-PAS)
select(climate, MAP, PAR, PAS)

#7 Lowest average spring monthly temperature
# Find lowest Tmin value between May and June
climate <- climate %>% rowwise() %>% mutate(Tmin_mj = min(c_across(Tmin_05:Tmin_06)))
select(climate, Tmin_mj)

#8 Lowest average fall monthly temperature (Tmin_at)
# Find the lowest Tmin value between August and September
climate <- climate %>% rowwise() %>% mutate(Tmin_as = min(c_across(Tmin_08:Tmin_09)))
select(climate, Tmin_as)

####Create clean dataset
#Pull out the variables described above
#Also include MAT (mean annual temperature), TD (temperature difference between mean coldest month temperature and mean warmest month temperature) and MAP (mean annual precipitation) because they are good descriptors
climate <- climate %>% 
  select(site_id, cluster, bec, MAT, TD, MAP, PAR, PAS, DD5_gs, Rad_gs, 
         Tmin_all, Tmin_sd, Tmin_mj, Tmin_as) #select the varaibles of interest
print(climate)

#Export this as a table for thesis
# write_csv(climate, "./tables/site_climate.csv")

####Graph this
#Create a long dataframe so a variables can be plotted in one go
# clim_g <- climate %>% 
#   pivot_longer(MAT:Tmin_as, names_to = "climvar", values_to = "climval")
# 
# #Create a plot
# ggplot(clim_g, aes(x=cluster, y=climval, label=site_id, color = bec)) + 
#   geom_text_repel() +
#   geom_point() + 
#   facet_wrap(~climvar, scales = "free_y") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, 
#                                    hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels
#   scale_x_discrete(labels = function(x) str_replace_all(x, " ", "\n")) #add a line break wherever there is a space
```

#SECTION 2: Size class distribution and species composition
Goal of this section is to compare the size class distribution and species composition in the mature and regen components between the sites. 
```{r}
#Calculate basal area for each tree in m^2 (m^2 is how its usually expressed at the tree level)
##https://en.wikipedia.org/wiki/Basal_area
trees <- trees %>% mutate(ba_m2 = pi*((dbh/100)/2)^2)

#Create bins of diameter classes from dbh
##Default is for intervals to be half-open. They include the upper bound but not the lower bound. e.g. (10,20] includes 20 but (20, 30] does not
## https://stackoverflow.com/questions/41304960/how-to-create-breaks-using-the-cut-function-without-numbers-overlapping
max(trees$dbh)
trees <- trees %>% mutate(dbh_bin = cut(dbh, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 
                                                        90, 100, 110, 120, 130, 140, 150, 
                                                        160, 170, 180, 190, 200), 
                                        include.lowest = TRUE))

#Create dataset thats comparable between sites. Need to do two things: 
##At a few sites we mapped mature trees up to 15m from the edge, when the standard was 10m. Need to remove extra trees mapped outside 10m. 
##Regen transects are variable length. Filter to just trees within 15m (the shortest transect) so that we are comparing the same transect footprint at each site
trees_comp <- trees %>% filter(outside_10 =="N" | is.na(outside_10)) %>%
  filter(dist_y_h<=15 | is.na(dist_y_h))

#Subdivide data into mature and regen components
##Mature
mature_comp <- trees_comp %>% filter(tree_type=="mature") #mature trees
##Regen
regen_15 <- trees_comp %>% filter(tree_type == "regen") #regen trees 
```

Compare the composition and size class distribution of the sites. 
Regen: 
-Sharp decreasing curve at: mk_3, mi_2, mk_1 and ph_1. Not completely explained by age or site history but these help. mi_1 was spaced at some point. Campbell river sites are in xm and typically have flatter distributions. 
-Not much to say on basal area. They represent sites at different points in a successional trajectory. 

Mature
-Maybe a BECsz effect. vm sites have more stems and peak in smaller dbh classes. 
-BA, large portions of the BA from larger diameter classes comes from Fd/Cw.
-Need to try to get better age estimates from mature portion.
```{r}
#Calculate some basic site level metrics: basal area, number of stems
##Regen
##Regen transect footprint = 15x5m=75m2=0.0075ha (but three transects/site) so 0.0075*3 = 0.0225ha 
reg.ba_nstem <- regen_15 %>% group_by(id_bec_age_f) %>% 
  summarise(reg.tot_ba_m2 = sum(ba_m2),  reg.nstem = n()) #calculate total BA and number of stems
reg.ba_nstem <- reg.ba_nstem %>% mutate(reg.tot_ba_m2ha = reg.tot_ba_m2/0.0225, 
                                     reg.nstem_ha = reg.nstem/0.0225) #Add a columns with per hectare numbers for easy reading

##Mature
##Mature component footprint = 55x10m =550m2=0.055ha. 
mat.ba_nstem <- mature_comp %>% group_by(id_bec_age_f) %>% 
  summarise(mat.tot_ba_m2 = sum(ba_m2),  mat.nstem = n()) #calculate total BA and number of stems
mat.ba_nstem <- mat.ba_nstem %>% mutate(mat.tot_ba_m2ha = mat.tot_ba_m2/0.055, 
                                     mat.nstem_ha = mat.nstem/0.055) #Add a columns with per hectare numbers for easy reading

#Join these to site table
site <- left_join(site, reg.ba_nstem, by="id_bec_age_f")
site <- left_join(site, mat.ba_nstem, by="id_bec_age_f")

#Create a summary table, see how they compare:
site %>% select(id_bec_age_f, reg.tot_ba_m2ha, reg.nstem_ha) %>% 
  arrange(reg.tot_ba_m2ha)
site %>% select(id_bec_age_f, mat.tot_ba_m2ha, mat.nstem_ha) %>% 
  arrange(mat.tot_ba_m2ha)
####################

#Create a summary of dbh at each site
##Calculate the 25th percentile, median and 75th percentile
reg.dbh <- regen_15 %>% group_by(id_bec_age_f) %>% 
  summarise(reg.dbh_25perc = quantile(dbh, probs = .25),
            reg.dbh_med = quantile(dbh, probs = .50),
            reg.dbh_75perc = quantile(dbh, probs = .75)) #calculate total BA and number of

mat.dbh <- mature_comp %>% group_by(id_bec_age_f) %>% 
  summarise(mat.dbh_25perc = quantile(dbh, probs = .25),
            mat.dbh_med = quantile(dbh, probs = .50),
            mat.dbh_75perc = quantile(dbh, probs = .75)) #calculate total BA and number of

#Do the same but just for Hw
reg.hw_dbh <- regen_15 %>% filter(spp=="Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(reg.dbh_hw_25perc = quantile(dbh, probs = .25),
            reg.dbh_hw_med = quantile(dbh, probs = .50),
            reg.dbh_hw_75perc = quantile(dbh, probs = .75)) #calculate total BA and number of

mat.hw_dbh <- mature_comp %>% filter(spp=="Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(mat.dbh_hw_25perc = quantile(dbh, probs = .25),
            mat.dbh_hw_med = quantile(dbh, probs = .50),
            mat.dbh_hw_75perc = quantile(dbh, probs = .75)) #calculate total BA and number of

#Join these back to the site dataframe
site <- left_join(site, reg.dbh, by="id_bec_age_f")
site <- left_join(site, mat.dbh, by="id_bec_age_f")
site <- left_join(site, reg.hw_dbh, by="id_bec_age_f")
site <- left_join(site, mat.hw_dbh, by="id_bec_age_f")
####################

#Now break down basal area and nstem, by species and diameter class.
##Bit of a messy table, but graphed below
reg.struc <- regen_15 %>% group_by(id_bec_age_f, spp, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% 
  complete(dbh_bin, fill=list(ba_m2=0, nstem=0)) #add 0s in places there was no trees
mat.struc <- mature_comp %>% group_by(id_bec_age_f, spp, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% 
  complete(dbh_bin, fill=list(ba_m2=0, nstem=0))#add 0s in places there was no trees

#Filter to the four dominant conifer species (hemlock (Hw) cedar (Cw), douglas fir (Fd) and amabilis fir (Ba)) to make comparisons easier.
##First check what percentage of the BA and # of stems Hw, Cw, Fd and Ba make up together (also add Hw alone because its focal species)
###Regen
reg.hwcwfdba_ba_m2 <- reg.struc %>% filter(spp %in% c("Hw", "Cw", "Fd", "Ba")) %>% group_by(id_bec_age_f) %>% 
  summarise(reg.hwcwfdba_ba_m2 = sum(ba_m2)) #calculate total BA of Hw, Cw, Fd and Ba at each site
reg.hwcwfdba_nstem <- reg.struc %>% filter(spp %in% c("Hw", "Cw", "Fd", "Ba")) %>% group_by(id_bec_age_f) %>% 
  summarise(reg.hwcwfdba_nstem = sum(nstem)) #calculate total # of stems of Hw, Cw, Fd and Ba at each site
reg.hw_ba_m2 <- reg.struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(reg.hw_ba_m2 = sum(ba_m2)) #calculate BA of Hw at each site
reg.hw_nstem <- reg.struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(reg.hw_nstem = sum(nstem)) #calculate number of Hw trees at each site

###Join variables back to site data
site <- left_join(site, reg.hwcwfdba_ba_m2, by="id_bec_age_f")
site <- left_join(site, reg.hwcwfdba_nstem, by="id_bec_age_f")
site <- left_join(site, reg.hw_ba_m2, by="id_bec_age_f")
site <- left_join(site, reg.hw_nstem, by="id_bec_age_f")
site <- site %>% mutate(reg.p_ba_hwcwfdba = reg.hwcwfdba_ba_m2/reg.tot_ba_m2, #proportion BA Hw, Fd, Cw Ba
                        reg.p_stem_hwcwfdba = reg.hwcwfdba_nstem/reg.nstem, #proportion stems Hw, Fd, Cw Ba
                        reg.p_ba_hw = reg.hw_ba_m2/reg.tot_ba_m2, #proportion BA Hw
                        reg.p_stem_hw = reg.hw_nstem/reg.nstem) #proportion stems Hw
###Mature
mat.hwcwfdba_ba_m2 <- mat.struc %>% filter(spp %in% c("Hw", "Cw", "Fd", "Ba")) %>% group_by(id_bec_age_f) %>% 
  summarise(mat.hwcwfdba_ba_m2 = sum(ba_m2)) #calculate total BA of just Hw, Cw, Fd and Ba
mat.hwcwfdba_nstem <- mat.struc %>% filter(spp %in% c("Hw", "Cw", "Fd", "Ba")) %>% group_by(id_bec_age_f) %>% 
  summarise(mat.hwcwfdba_nstem = sum(nstem)) #calculate total # of stems of Hw, Cw, Fd and Ba at each site
mat.hw_ba_m2 <- mat.struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(mat.hw_ba_m2 = sum(ba_m2)) #calculate BA of Hw
mat.hw_nstem <- mat.struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(mat.hw_nstem = sum(nstem)) #calculate number of Hw trees at each site

###Join variables back to site data
site <- left_join(site, mat.hwcwfdba_ba_m2, by="id_bec_age_f")
site <- left_join(site, mat.hwcwfdba_nstem, by="id_bec_age_f")
site <- left_join(site, mat.hw_ba_m2, by="id_bec_age_f")
site <- left_join(site, mat.hw_nstem, by="id_bec_age_f")
site <- site %>% mutate(mat.p_ba_hwcwfdba = mat.hwcwfdba_ba_m2/mat.tot_ba_m2, #proportion BA Hw, Fd, Cw Ba
                        mat.p_stem_hwcwfdba = mat.hwcwfdba_nstem/mat.nstem, #proportion stems Hw, Fd, Cw Ba
                        mat.p_ba_hw = mat.hw_ba_m2/mat.tot_ba_m2, #proportion BA Hw
                        mat.p_stem_hw = mat.hw_nstem/mat.nstem) #proportion stems Hw

##Create summary table. Hw, Cw, Fd and Ba make up between 78-100% of regen basal area, and 88-100% of mature basal area. So only visualizing them seems reasonable.
## Hw makes up between 12 and 95% in regen and between 10% and 97% in mature component. Two Ucluelet sites (ph_2 and ph_3) are outliers, with low Hw percentages. Spread of Hw percentage more continuous in mature; in regen, most sites have >70% Hw and only ph_2 and ph_3 are <50%. 
site %>% select(id_bec_age_f, reg.tot_ba_m2, reg.hwcwfdba_ba_m2, reg.hw_ba_m2, reg.p_ba_hwcwfdba, reg.p_ba_hw) %>% 
  arrange(reg.tot_ba_m2)
site %>% select(id_bec_age_f, mat.tot_ba_m2, mat.hwcwfdba_ba_m2, mat.hw_ba_m2, mat.p_ba_hwcwfdba, mat.p_ba_hw) %>% 
  arrange(mat.tot_ba_m2)

##Filter
reg.struc <- filter(reg.struc, spp %in% c("Hw", "Cw", "Fd", "Ba"))
mat.struc <- filter(mat.struc, spp %in% c("Hw", "Cw", "Fd", "Ba"))

#Make some plots:
levels(reg.struc$spp)
##Regen
# reg.struc <- reg.struc %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
# ggplot(reg.struc, aes(x=dbh_bin, y=nstem, fill=spp)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)
# ggplot(reg.struc, aes(x=dbh_bin, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)

##Mature
# mat.struc <- mat.struc %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
# ggplot(mat.struc, aes(x=dbh_bin, y=nstem, fill=spp)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)
# ggplot(mat.struc, aes(x=dbh_bin, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)

##############

#Assemble summary table for thesis
##Pull together variables
comp_size <-
  select(
    .data = site, 
    site_id,
    reg.tot_ba_m2,
    mat.tot_ba_m2,
    reg.tot_ba_m2ha,
    mat.tot_ba_m2ha,
    reg.nstem,
    mat.nstem,
    reg.nstem_ha,
    mat.nstem_ha,
    reg.hw_ba_m2,
    mat.hw_ba_m2,
    reg.p_ba_hw,
    mat.p_ba_hw,
    reg.hw_nstem,
    mat.hw_nstem,
    reg.p_stem_hw,
    mat.p_stem_hw,
    reg.dbh_25perc,
    mat.dbh_25perc,
    reg.dbh_med,
    mat.dbh_med,
    reg.dbh_75perc,
    mat.dbh_75perc,
    reg.dbh_hw_25perc,
    mat.dbh_hw_25perc,
    reg.dbh_hw_med,
    mat.dbh_hw_med,
    reg.dbh_hw_75perc,
    mat.dbh_hw_75perc
  )

#Pivot the table longer so every measured variable has one column and two rows, one for mature trees and one for regen trees
comp_size <- comp_size %>% pivot_longer(cols = -site_id, names_to = c("tree_type", ".value"), names_sep = "\\.")
comp_size <- comp_size %>% mutate(tree_type = case_when(tree_type == "reg" ~ "regen", #recode tree_type
                                                        tree_type == "mat" ~ "mature"))

#Export table
#Export this as a table for thesis
# write_csv(comp_size, "./tables/comp_size.csv")
```
#SECTION 3: Compare infection levels in mature trees
Compare the HDM infection in the mature component across the sites by looking at: basal area and stem count distributions parsed out by presence/absence of hdm infection and dmr. Generally trying to assess whether the mature components are similar infection sources. 
- Some spread in the proportion of the Hw basal area infected (37% at mk_1 to 100% at cr_1) but 9/10 sites>70 and 8/10 > 80%.
- Mature components have different basal areas so its important to look at the magnitude (not just proportion) infected trees
- Graph comparing dmr across diameter classes is super useful. Sites with old (big) mature components with high levels of infection generate greatest "inoculum load". 
```{r}
#Filter to just Hw and Ba trees because these are the only ones we are considering "significant hosts"
mat.hwba <- mature_comp %>% filter(spp %in% c("Hw", "Ba"))
mat.hwba_summary <- mat.hwba %>% group_by(id_bec_age_f) %>% 
  summarise(mat.hwba_ba_m2 = sum (ba_m2), mat.hwba_nstem = n()) #calculate summary statistics
site <- left_join(site, mat.hwba_summary, by="id_bec_age_f") #add summary stats back to site data

#Start simple and calculate the basal area and number of infected trees at each site
mat.inf <- mat.hwba %>% filter(hdm_pa =="Y") %>% group_by(id_bec_age_f) %>% 
  summarise(mat.inf_ba_m2 = sum(ba_m2), mat.inf_nstem = n()) #infected
site <- left_join(site, mat.inf, by="id_bec_age_f") #join variables back to site data
site <- site %>% mutate(mat.p_hwba_ba_inf = mat.inf_ba_m2/mat.hwba_ba_m2, 
                        mat.p_hwba_nstem_inf = mat.inf_nstem/mat.hwba_nstem) #calculate the proportion of signifcant hosts (Hw and Ba) stems and basal area that are infected at each site

#Create a summary table. 
site %>% select(id_bec_age_f, mat.tot_ba_m2, mat.hwba_ba_m2, mat.inf_ba_m2, mat.p_hwba_ba_inf) %>% #summary table sorted by proportion of BA infected
  arrange(mat.p_hwba_ba_inf) 
site %>% select(id_bec_age_f, mat.nstem, mat.hwba_nstem, mat.inf_nstem, mat.p_hwba_nstem_inf) %>% #summary table sorted by proportion of BA infected
  arrange(mat.p_hwba_nstem_inf) 

#Now calculate basal area and # of stems by dmr
mat.dmr <- mat.hwba %>% group_by(site_id, dmr_f2) %>% summarise(ba_m2 = sum(ba_m2), nstem =n()) %>% 
  complete(dmr_f2, fill=list(ba_m2=0, nstem=0))
mat.dmr <- mat.dmr %>% filter(dmr_f2 != "-")
mat.dmr <- pivot_wider(mat.dmr, names_from = dmr_f2, values_from = c(ba_m2, nstem))

#Create table for thesis
mat_infection <- site %>% select(site_id, mat.tot_ba_m2, mat.nstem, mat.inf_ba_m2, mat.inf_nstem, mat.p_hwba_ba_inf,
                                 mat.p_hwba_nstem_inf)
mat_infection <- left_join(mat_infection, mat.dmr, by = "site_id")

#Export this for thesis
# write_csv(mat_infection, "./tables/mat_infection.csv")
# 
# #Now graph infection and DMR across dbh distributions
# ##Break down basal area and # stems by site, presence/absece of hdm and diameter class
# mat.hdmpa_dbh <- mat.hwba %>% group_by(id_bec_age_f, hdm_pa, dbh_bin) %>% 
#   summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% complete(dbh_bin, fill=list(ba_m2=0, nstem=0))
# 
# ##Graph this across diameter classes
# mat.hdmpa_dbh <- mat.hdmpa_dbh %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
# ggplot(mat.hdmpa_dbh, aes(x=dbh_bin, y=nstem, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
# ggplot(mat.hdmpa_dbh, aes(x=dbh_bin, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
# 
# ##Do the same but by DMR
# mat.dmr_dbh <- mat.hwba %>% group_by(id_bec_age_f, dmr_f2, dbh_bin) %>% 
#   summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% complete(dbh_bin, fill=list(ba_m2=0, nstem=0))
# 
# ##Graph this across diameter classes
# mat.dmr_dbh <- mat.dmr_dbh %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
# ggplot(mat.dmr_dbh, aes(x=dbh_bin, y=nstem, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
# ggplot(mat.dmr_dbh, aes(x=dbh_bin, y=ba_m2, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
#   facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
```

#SECTION 4 - Compare infection levels and distribution of infection in the regen compoenent 
- Note variable length transects were used and we need to account for this when comparing between sites. Calculating the proportion of the total basal area that is infected at a site isn't useful. i.e. 50% infection at site 1 vs 40% infection at site 2 isn't a meaningful comparison if site 1 and site 2 have different length transects. Deal with this in two ways: compare total magnitude (e.g. basal area infected) --> sites with longer transects are places where more spread has occured amd will likely have higher total magnitudes OR compare discrete intervals (e.g. basal area infected within 15m)

- Infection perimeter:
  - Range: 2-23m, but with signifcant variability between transects within a site
  - For the most part, it is coupled with basal area infected. Plot 1 (p.infper_ba) shows that for all sites except ph_1, sites with     high basal area also have larger infection perimeter
  
-Proportion Hw infected within 15m
  -by # of stems, range: 1-82%, mean = 47%
  -by basal area, range: .5-92%, mean = 57%
  -four outliers are obvious in this data: mk_3 (youngest site by far), ph_1 (no clear explanation), ph_2 and ph_3 (both sites where     Hw is not a dominant species)
  
-DMR
  -Generally, low, as you would expect in regen
  -Stand DMR ranges from 0-1.6, mean = .6
  -Stand DMI ranges from 0-2.1, mean = 1.233
  -Plot 2 visualizes this. Most Hw are uninfected, those that are infected fall into the IBLC and 1-2 dmr range. The plots that have regen trees in the dmr-56 category, it would be worth looking at which were cored and identified as advanced regen/residuals.
  
-Plots relating the propotion of infected stems, proportion of infected basal area and dmr distribution to distance from the edge
  -This is plot 3
  -The amount Hw is variable as a function of distance from the edge and the infection metrics seem to follow it (makes sense, bigger target area)
  - There are patterns of percent infection and dmr descreasing as a function of distance from the edge at some sites (e.g. cr_1, cr_2, mk_2, mi_1)
  - The severely infected trees (dmr 5-6) are all close to the edge (within 10m)

-Plots relating infection in regen to basal area of infection in mature component
  -This is Plot 4 (which combines p.minfba_rinfba and p.minfba_rpinfba)
  -There is a positive relationship here but only if you mentally remove the outliers (particularly mk_3 and ph_1)

-Plots relating infection perimeter to regen age
  -This is Plot 5
  -It sort of looks like there is something here but my sense is the range of age is too small and the replication is too low
```{r}
#Create dataset of just Hw regen trees (our species of interest)
reg.hw <- trees %>% filter(tree_type == "regen" & spp=="Hw") #Hemlock regen trees 

#Calculate the infection perimeter at each site
##Use the 90th percentile to do this. This metric is: the distance from the edge below which 90% of the infected Hw trees occur, averaged across the three transects at a site
reg.infper <- reg.hw %>% group_by(id_bec_age_f, plot_id, hdm_pa) %>% 
  summarise(reg.infper = quantile(dist_y_h, probs = 0.9)) %>% 
  complete(hdm_pa, fill=list(reg.infper=0)) #get infection perimeter for each transect, grouped by hdm_pa
reg.infper <- reg.infper %>% filter(hdm_pa == "Y") #filter to infected trees, can't filter first because we want to keep transects with no infection (i.e 0 is a valid value) 
range(reg.infper$reg.infper) #0's kept
reg.infper <- reg.infper %>% group_by(id_bec_age_f) %>% 
  summarise(reg.infper_m = mean(reg.infper), 
            reg.infper_sd = sd(reg.infper)) #get site level estimates of mean infection perimeter and its standard deviation
##Join these back to the site dataframe
site <- left_join(site, reg.infper, by="id_bec_age_f")

#Then calculate basal area and number of stems infected at each site
reg.inf <- reg.hw %>% filter(hdm_pa =="Y") %>% 
  group_by(id_bec_age_f) %>% summarise(reg.inf_ba_m2 = sum(ba_m2), reg.inf_nstem = n()) #infected Hw, full transect length
reg.15_hw <- regen_15 %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(reg.15_hw_ba_m2 = sum(ba_m2), reg.15_hw_nstem = n()) #all Hw, first 15m
reg.15_inf <- regen_15 %>% filter(spp == "Hw" & hdm_pa =="Y") %>% 
  group_by(id_bec_age_f) %>% summarise(reg.15_inf_ba_m2 = sum(ba_m2), reg.15_inf_nstem = n()) #infected Hw, first 15m
site <- left_join(site, reg.inf, by="id_bec_age_f") #join variables back to site data
site <- left_join(site, reg.15_hw, by="id_bec_age_f")
site <- left_join(site, reg.15_inf, by="id_bec_age_f")
site <- site %>% mutate(reg.p_hw15_ba_inf = reg.15_inf_ba_m2/reg.15_hw_ba_m2,
                        reg.p_hw15_nstem_inf = reg.15_inf_nstem/reg.15_hw_nstem)

#Now look at the distribution of infection by dmr rating
##Start by computing stand dmr and dmi for each site
##stand dmr = average dmr for all susceptible trees, defined here as Hw trees (excluding Ba because they only occur on one site)
##note: the way these are calculated, dead trees are excluded (whether they are infected or not) and trees infected only below live crown are considered dmr 0 
##stand dmi = average dmr of all infected susceptible trees, a measure of severity
##note: similarly, infected dead trees are excluded) and trees infected only below live crown are considered dmr 0 
reg.s_dmr <- reg.hw %>% group_by(site_id) %>% summarise(reg.s_dmr = mean(dmr, na.rm = T))
reg.s_dmi <- reg.hw %>% filter(status %in% c("LS", "LL", "LF") & hdm_pa == "Y") %>% 
  group_by(site_id) %>% summarise(reg.s_dmi = mean(dmr, na.rm = T))

##Join these back to the site table
site <- left_join(site, reg.s_dmr, by="site_id")
site <- left_join(site, reg.s_dmi, by="site_id")

#Now find the basal area and number of stems in each dmr class
reg.dmr <- reg.hw %>% group_by(site_id, dmr_f2) %>% summarise(ba_m2 = sum(ba_m2), nstem =n()) %>% 
  complete(dmr_f2, fill=list(ba_m2=0, nstem=0))
reg.dmr <- reg.dmr %>% filter(dmr_f2 != "-")
reg.dmr <- pivot_wider(reg.dmr, names_from = dmr_f2, values_from = c(ba_m2, nstem))

#Create a summary table and export it for thesis
reg_infection <- site %>% select(site_id, reg.infper_m, reg.infper_sd, reg.tot_ba_m2, reg.nstem, reg.hw_ba_m2, reg.hw_nstem, 
                                  reg.15_hw_ba_m2, reg.15_hw_nstem, reg.inf_ba_m2, reg.inf_nstem, 
                                  reg.15_inf_ba_m2, reg.p_hw15_ba_inf, reg.15_inf_nstem, reg.p_hw15_nstem_inf, reg.s_dmr,
                                  reg.s_dmi)
reg_infection <- left_join(reg_infection, reg.dmr, by="site_id")

#write_csv(reg_infection, "./tables/reg_infection.csv")

#Create some exploratory graphs of the regen level infection metrics
##Plot set 1: dual axis plot looking at infection perimeter and basal area infected
##Need to set a scale factor for second axis. basal area about 24 times smaller than infper values, so times by 24
scale <- 24
p.infper_ba <- ggplot(site, aes(x=site_id, y=reg.infper_m)) + 
  geom_point(aes(color = "infection perimeter (m2)")) +
  geom_point(aes(y=reg.inf_ba_m2*scale, color = "basal area infected (m2)")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="basal area infected (m2)"))

#Plot set 2: basic distribution of dmr by site
g.reg_dmr <- reg.hw %>% group_by(site_id, dmr_f2) %>% summarise(ba_m2 = sum(ba_m2), nstem =n()) %>% 
  complete(dmr_f2, fill=list(ba_m2=0, nstem=0))
g.reg_dmr <- g.reg_dmr %>% filter(dmr_f2 != "-")
p.reg_dmr_ba <- ggplot(g.reg_dmr, aes(x=dmr_f2, y=ba_m2, fill = dmr_f2)) + geom_bar(position="stack", stat="identity") +
  facet_wrap(vars(site_id)) + scale_fill_manual(values = colors_dmr_f2)
p.reg_dmr_nstem <- ggplot(g.reg_dmr, aes(x=dmr_f2, y=nstem, fill = dmr_f2)) + geom_bar(position="stack", stat="identity") +
  facet_wrap(vars(site_id)) + scale_fill_manual(values = colors_dmr_f2)

#Plot set 3: basal area and number of stems infected as function of distance from the edge
##Break up dist_y_h into intervals. Need these to make barcharts
summary(reg.hw$dist_y_h)
reg.hw <- reg.hw %>% mutate(dist_y_h_bin = cut(dist_y_h, breaks = c(0, 5, 10, 15, 20, 25, 30, 35), include.lowest = TRUE))
levels(reg.hw$dist_y_h_bin)

##Summarize basal area and the # stems by site, presence/absence of hdm and distance from the edge (in the 5m intervals created above)
reg.hdmpa_disty <- reg.hw %>% group_by(id_bec_age_f, hdm_pa, crown_class, dist_y_h_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% 
  complete(dist_y_h_bin, fill=list(ba_m2=0, nstem=0))

##Plot nstem and basal area vs. distance from the edge, coloured by presence/absence of hdm infection
ggplot(reg.hdmpa_disty, aes(x=dist_y_h_bin, y=nstem, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_hdm_pa)
ggplot(reg.hdmpa_disty, aes(x=dist_y_h_bin, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_hdm_pa)

##Do the same but with DMR. 
reg.dmr_disty <- reg.hw %>% group_by(id_bec_age_f, dmr_f2, crown_class, dist_y_h_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), nstem = n()) %>% 
  complete(dist_y_h_bin, fill=list(ba_m2=0, nstem=0))

##Plot nstem and basal area across distance from the edge, coloured by dmr
ggplot(reg.dmr_disty, aes(x=dist_y_h_bin, y=nstem, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
ggplot(reg.dmr_disty, aes(x=dist_y_h_bin, y=ba_m2, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)

#Plot set 4: basal area infected in mature component to basal area infected in the regen
p.minfba_rinfba <- ggplot(site, aes(x=mat.inf_ba_m2, y=reg.inf_ba_m2, color=id_bec_age_f)) + 
  geom_point()
p.minfba_rpinfba <- ggplot(site, aes(x=mat.inf_ba_m2, y=reg.p_hw15_ba_inf, color=id_bec_age_f)) + 
  geom_point()
print(p.minfba_rinfba/p.minfba_rpinfba)

#Plot set 5: infection perimeter as a function of site age
ggplot(site, aes(x=age, y=reg.infper_m, color=id_bec_age_f)) + 
  geom_point()
```

Extra code
```{r}
####################################
#Start to build a dataframe at the transect level for plotting
tran <- trees %>% filter(tree_type=="regen") %>% select(site_id, plot_id) %>% 
  distinct() #step 1: two columns, site_id and plot_id
tran <- left_join(tran, select(site, age, cluster, bec_sz, site_id), 
                  by="site_id") #step 2: add site level variables
#step 3: calculate some metrics of spread
#metric 1: 75% percentiles of infected trees for each transect (i.e. the distance where 75% of the infected trees are between there and the mature component)
#metric 2: basal area infected
spread_metrics <- regen %>% filter(hdm_pa == "Y") %>% group_by(plot_id) %>% 
  summarise(q75 = quantile(dist_y_h, probs = c(0.75)), ba_inf=sum(ba_m2))
ba_15_inf <- regen %>% filter(hdm_pa == "Y" & dist_y_h<15) %>% group_by(plot_id) %>% 
  summarise(ba_15_inf = sum(ba_m2))
ba_15_tot <- regen %>% filter(dist_y_h<15) %>% group_by(plot_id) %>% summarise(ba_15_tot = sum(ba_m2))
tran <- left_join(tran, spread_metrics, by="plot_id")
tran <- left_join(tran, ba_15_inf, by="plot_id")
tran <- left_join(tran, ba_15_tot, by="plot_id")
tran <- tran %>% replace_na(list(q75 = 0, ba_inf = 0, ba_15_inf=0)) #replace NAs with 0
tran <- tran %>% mutate(p_ba_15_inf = ba_15_inf/ba_15_tot)

#Scatterplots
ggplot(tran, aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age")
ggplot(filter(tran, site_id!="mk_3"), aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age, minus outlier")
ggplot(tran, aes(x=age, y=ba_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age")
ggplot(filter(tran, site_id!="mk_3"), aes(x=age, y=ba_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age, minus outlier")
ggplot(tran, aes(x=age, y=p_ba_15_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age, minus outlier")
```

