---
title: "Site Level Analyses"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
WHAT?
Script for running site level analyses in the general form (spread measure ~ site variable).

UPDATED: 28 Jun 2024

```{r}
rm(list=ls(all=TRUE))

#Load packages
library(tidyverse)
library(RColorBrewer)
library(ggrepel)
```

#DATA PREP

Read in and format the site and tree datasets
```{r}
####Read in data and inspect it
##Dataset 1: site data. 
#This contains variables at the site level. The key pieces for analysis are: year harvested and CWH subzone
site <- read_csv("data/cleaned/site data.csv")
str(site)
summary(site)

##Create some variables to use in summary tables and figures
#Calculate approximate age of regen. Sites assessed in 2023 so age = 2023 - year harvested
site <- site %>% mutate(age = 2023-yr_har)
site$age #luckily, got pretty good spread

#Create a single variable from all the pieces of a BEC zone
site <- site %>% unite(col = bec, sep = "", c("bec_z", "bec_sz", "bec_var"), remove = FALSE)

#Create a variable that combines site_id, bec, and regen age 
site <- site %>% unite(col = id_bec_age_f, sep = ", ", 
                       c("site_id", "bec", "age"), 
                       remove = FALSE) %>% 
  mutate(id_bec_age_f = factor(id_bec_age_f))

#Make cluster, bec and id_bec_age_f factors
site <- site %>% mutate(across(c(cluster, bec, id_bec_age_f), ~as.factor(.)))
str(site)

##Dataset 2: tree data
#This contains data for each measured tree (mature and regen trees). 
#Treats assessed_by var as a logical for some reason and throws up an error. Not a problem and not going to use this var
trees <- read_csv("data/workflow/trees_mapped.csv")
str(trees)
summary(trees)

#Convert factor vars to factors: 
trees <- trees %>% mutate(across(
  c(site_id, spp, status, hdm_pa, b_lc, 
    broom_pa, broom_pos, stem_pa, crown_class, crown_cond, outside_10, assessed_by, tree_type, dmr_f),
  ~as.factor(.)))

#Convert plot_id and dmrs to integers
trees <- trees %>% mutate(across(
  c(plot_id, dmr_l, dmr_m, dmr_u, dmr), ~as.integer(.)))

str(trees)

#Create new factor dmr variable that has fewer levels and ascends logically from healthy to most infected
levels(trees$dmr_f)
trees <- trees %>% mutate(dmr_f2 = case_when((dmr_f =="1" | dmr_f=="2") ~ "1-2", (dmr_f =="3" | dmr_f=="4") ~ "3-4", (dmr_f =="5" | 
                                    dmr_f=="6") ~ "5-6", .default = dmr_f))
trees <- trees %>% mutate(dmr_f2 = factor(dmr_f2, levels = c("-", "DU", "0","IBLC", "1-2", "3-4", "5-6", "DI")))
summary(trees$dmr_f2)

#Join some site level variables to tree data: 
trees <- left_join(trees, select(site, cluster, 
                                 bec, id_bec_age_f, age, site_id), by="site_id")
```

Define some consistent colours to plot with
```{r}
#spp colours Cw, Fd, Hw
colors_spp = c("firebrick", "burlywood4", "dodgerblue")

#hdm_pa colours. Didn't include a colour for non-Hw trees because not creating any plots with them and this variable here. 
levels(trees$hdm_pa)
colors_hdm_pa <- c("dodgerblue", "grey", "goldenrod1")

#dmr_f2 colours. Didn't include a colour for non-Hw trees because not creating any plots with them and this variable here. 
levels(trees$dmr_f2)
colors_dmr_f2 = c("dodgerblue4", "dodgerblue", "khaki1", "khaki3", "goldenrod1", "goldenrod3", "hotpink")
```


#SECTION 1: Site climate comparisons
Goal of this section is to descriptively describe differences in climate variables between sites that might influence HDM biology.
```{r}
####Read in climate data
#Data downloaded from ClimateBC (https://climatebc.ca/mapVersion, 24 Jun 2024). ClimateBCv7.50. Period: Normal_1991_2020
climate <- read_csv('./data/raw/hdm_climdata.csv') #raw data from climateBC

####Check coordinates were inputted correctly, then join site_ids 
#Sites are identified by coordinates in the climate dataset. Check they are right before joining site_ids
site <- site %>% mutate(dd_X_r = round(dd_X, digits = 4), #round number of decimals to match climateBC rounding
                        dd_Y_r = round(dd_Y, digits = 5))
x <- site$dd_X_r - climate$long
y <- site$dd_Y_r - climate$Lat
print(x); print(y)
#Good, almost al 0s, just one that is different by 1 x 10e-4 (a rounding error)

#Join site_id, bec and cluster by latitude values
climate <- left_join(climate, select(site, site_id, cluster, bec, dd_Y_r), by = c("Lat" = "dd_Y_r"))

#move site_id to the front
climate <- climate %>% select(site_id, cluster, bec, everything())

#### Define climate variables that were in Bianca's paper
#Descriptions of these variables and whether they are exact matches or approximiations of variables used in paper are in: /Users/hannosoutham/Library/CloudStorage/OneDrive-UBC(1)/Msc/Thesis/ch1.docx

#1 Growing degree days
climate <- climate %>% rowwise() %>% mutate(DD5_gs = sum(c_across(DD5_sp:DD5_at)))
select(climate, DD5, DD5_gs)

#2 Growing season radians
climate <- climate %>% rowwise() %>% mutate(Rad_gs = mean(c_across(Rad_sp:Rad_at)))
select(climate, MAR, Rad_gs)

#3 Lowest average minimum monthly temperature (Tmin_wt)
# find lowest Tmin value across all months
climate <- climate %>% rowwise() %>% mutate(Tmin_all = min(c_across(Tmin_01:Tmin_12)))
select(climate, Tmin_wt, Tmin_all)

#4 Standard deviation of lowest average minimum monthly temperature  
# calculate the standard deviation in Tmin across all of the months
climate <- climate %>% rowwise() %>% mutate(Tmin_sd = sd(c_across(Tmin_01:Tmin_12)))
select(climate, Tmin_sd)

#5 Snow (PAS) 
select(climate, MAP, PAS)

#6 Rain
climate <- climate %>% mutate(PAR = MAP-PAS)
select(climate, MAP, PAR, PAS)

#7 Lowest average spring monthly temperature
# Find lowest Tmin value between May and June
climate <- climate %>% rowwise() %>% mutate(Tmin_mj = min(c_across(Tmin_05:Tmin_06)))
select(climate, Tmin_mj)

#8 Lowest average fall monthly temperature (Tmin_at)
# Find the lowest Tmin value between August and September
climate <- climate %>% rowwise() %>% mutate(Tmin_as = min(c_across(Tmin_08:Tmin_09)))
select(climate, Tmin_as)

####Create clean dataset
#Pull out the variables described above
#Also include MAT (mean annual temperature), TD (temperature difference between mean coldest month temperature and mean warmest month temperature) and MAP (mean annual precipitation) because they are good descriptors
climate <- climate %>% 
  select(site_id, cluster, bec, MAT, TD, MAP, PAR, PAS, DD5_gs, Rad_gs, 
         Tmin_all, Tmin_sd, Tmin_mj, Tmin_as) #select the varaibles of interest
print(climate)

#Export this as a table for thesis
write_csv(climate, "./tables/site_climate.csv")

####Graph this
#Create a long dataframe so a variables can be plotted in one go
clim_g <- climate %>% 
  pivot_longer(MAT:Tmin_as, names_to = "climvar", values_to = "climval")

#Create a plot
ggplot(clim_g, aes(x=cluster, y=climval, label=site_id, color = bec)) + 
  geom_text_repel() +
  geom_point() + 
  facet_wrap(~climvar, scales = "free_y") +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1, vjust = 0.5)) +  # Rotate x-axis labels
  scale_x_discrete(labels = function(x) str_replace_all(x, " ", "\n")) #add a line break wherever there is a space
```

#SECTION 2: Size class distribution and species composition
Goal of this section is to compare the size class distribution and species composition in the mature and regen components between the sites. 

Calculate basal area and create a dataset that is comparable between sites. 
```{r}
#Calculate basal area for each tree in m^2 (m^2 is how its usually expressed at the tree level)
##https://en.wikipedia.org/wiki/Basal_area
trees <- trees %>% mutate(ba_m2 = pi*((dbh/100)/2)^2)

#Create bins of diameter classes from dbh
##Default is for intervals to be half-open. They include the upper bound but not the lower bound. e.g. (10,20] includes 20 but (20, 30] does not
## https://stackoverflow.com/questions/41304960/how-to-create-breaks-using-the-cut-function-without-numbers-overlapping
trees <- trees %>% mutate(dbh_bin = cut(dbh, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 
                                               110, 120, 130, 140, 150, 160, 170, 180), include.lowest = TRUE))

#Create dataset thats comparable between sites. Need to do two things: 
##At a few sites we mapped mature trees up to 15m from the edge, when the standard was 10m. Need to remove extra trees mapped outside 10m. 
##Regen transects are variable length. Filter to just trees within 15m (the shortest transect) so that we are comparing the same transect footprint at each site
trees_comp <- trees %>% filter(outside_10 =="N" | is.na(outside_10)) %>%
  filter(dist_y_h<=15 | is.na(dist_y_h))

#Subdivide data into mature and regen components
##Mature
mature_comp <- trees_comp %>% filter(tree_type=="mature") #mature trees
##Regen
regen_15 <- trees_comp %>% filter(tree_type == "regen") #regen trees 
```

Compare the composition and size class distribution of the sites. 
Regen: 
-Sharp decreasing curve at: mk_3, mi_2, mk_1 and ph_1. Not completely explained by age or site history but these help. mi_1 was spaced at some point. Campbell river sites are in xm and typically have flatter distributions. 
-Not much to say on basal area. They represent sites at different points in a successional trajectory. 

Mature
-Maybe a BECsz effect. vm sites have more stems and peak in smaller dbh classes. 
-BA, large portions of the BA from larger diameter classes comes from Fd/Cw.
-Need to try to get better age estimates from mature portion.
```{r}
#Calculate some basic site level metrics: basal area, number of stems
##Regen
##Regen transect footprint = 15x5m=75m2=0.0075ha (but three transects/site) so 0.0075*3 = 0.0225ha 
reg_ba_nstem <- regen_15 %>% group_by(id_bec_age_f) %>% 
  summarise(reg_tot_ba_m2 = sum(ba_m2),  reg_nstem = n()) #calculate total BA and number of stems
reg_ba_nstem <- reg_ba_nstem %>% mutate(reg_tot_ba_m2ha = reg_tot_ba_m2/0.0225, #Add a columns with per hectare numbers for easy reading
                                     reg_nstem_ha = reg_nstem/0.0225) 
##Mature
##Mature component footprint = 55x10m =550m2=0.055ha. 
mat_ba_nstem <- mature_comp %>% group_by(id_bec_age_f) %>% 
  summarise(mat_tot_ba_m2 = sum(ba_m2),  mat_nstem = n()) #calculate total BA and number of stems
mat_ba_nstem <- mat_ba_nstem %>% mutate(mat_tot_ba_m2ha = mat_tot_ba_m2/0.055, #Add a columns with per hectare numbers for easy reading
                                     mat_nstem_ha = mat_nstem/0.055) 

#Join these to site table
site <- left_join(site, reg_ba_nstem, by="id_bec_age_f")
site <- left_join(site, mat_ba_nstem, by="id_bec_age_f")

#Create a summary table, see how they compare:
site %>% select(id_bec_age_f, reg_tot_ba_m2ha, reg_nstem_ha) %>% 
  arrange(reg_tot_ba_m2ha)
site %>% select(id_bec_age_f, mat_tot_ba_m2ha, mat_nstem_ha) %>% 
  arrange(mat_tot_ba_m2ha)
####################

#Now break down basal area and nstems, by species and diameter class.
##Bit of a messy table, but graphed below
reg_struc <- regen_15 %>% group_by(id_bec_age_f, spp, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% 
  complete(dbh_bin, fill=list(ba_m2=0, n_stems=0)) #add 0s in places there was no trees
mat_struc <- mature_comp %>% group_by(id_bec_age_f, spp, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% 
  complete(dbh_bin, fill=list(ba_m2=0, n_stems=0))#add 0s in places there was no trees

#Filter to the three dominant conifer species to make comparisons easier.
##First check how much BA Hw, Cw and Fd make up together (also add Hw alone because its focal species)
###Regen
reg_hwcwfd_ba_m2 <- reg_struc %>% filter(spp %in% c("Hw", "Cw", "Fd")) %>% group_by(id_bec_age_f) %>% 
  summarise(reg_hwcwfd_ba_m2 = sum(ba_m2)) #calculate total BA of just Hw, Cw, Fd at each site
reg_hw_ba_m2 <- reg_struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(reg_hw_ba_m2 = sum(ba_m2)) #calculate BA of Hw at each site

###Join variables back to site data
site <- left_join(site, reg_hwcwfd_ba_m2, by="id_bec_age_f")
site <- left_join(site, reg_hw_ba_m2, by="id_bec_age_f")
site <- site %>% mutate(p_reg_ba_hwcwfd = reg_hwcwfd_ba_m2/reg_tot_ba_m2, 
                        p_reg_ba_hw = reg_hw_ba_m2/reg_tot_ba_m2) #calculate proportion of ba made up by Hw, Cw, Fd and then Hw alone

###Mature
mat_tot_ba_m2 <- mat_struc %>% group_by(id_bec_age_f) %>% 
  summarise(mat_tot_ba_m2 = sum(ba_m2)) #calculate total BA of maten
mat_hwcwfd_ba_m2 <- mat_struc %>% filter(spp %in% c("Hw", "Cw", "Fd")) %>% group_by(id_bec_age_f) %>% 
  summarise(mat_hwcwfd_ba_m2 = sum(ba_m2)) #calculate total BA of just Hw, Cw, Fd
mat_hw_ba_m2 <- mat_struc %>% filter(spp == "Hw") %>% group_by(id_bec_age_f) %>% 
  summarise(mat_hw_ba_m2 = sum(ba_m2)) #calculate BA of Hw

###Join variables back to site data
site <- left_join(site, mat_hwcwfd_ba_m2, by="id_bec_age_f")
site <- left_join(site, mat_hw_ba_m2, by="id_bec_age_f")
site <- site %>% mutate(p_mat_ba_hwcwfd = mat_hwcwfd_ba_m2/mat_tot_ba_m2,
                         p_mat_ba_hw = mat_hw_ba_m2/mat_tot_ba_m2) #calculate proportion of ba made up by Hw, Cw, Fd and then Hw alone

##Create summary table. Hw, Cw, Fd make up between 78-100% of regen basal area, and 88-100% of mature basal area. So filtering seems reasonable.
## Hw makes up between 66 and 95% in regen and 38% and 97% in mature component. Mature component interesting - not the same inoculum pressure.
site %>% select(id_bec_age_f, reg_tot_ba_m2, reg_hwcwfd_ba_m2, reg_hw_ba_m2, p_reg_ba_hwcwfd, p_reg_ba_hw) %>% 
  arrange(reg_tot_ba_m2)
site %>% select(id_bec_age_f, mat_tot_ba_m2, mat_hwcwfd_ba_m2, mat_hw_ba_m2, p_mat_ba_hwcwfd, p_mat_ba_hw) %>% 
  arrange(mat_tot_ba_m2)

##Filter
reg_struc <- filter(reg_struc, spp %in% c("Hw", "Cw", "Fd"))
mat_struc <- filter(mat_struc, spp %in% c("Hw", "Cw", "Fd"))

#Make some plots:
levels(reg_struc$spp)
##Regen
reg_struc <- reg_struc %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
ggplot(reg_struc, aes(x=dbh_bin, y=n_stems, fill=spp)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)
ggplot(reg_struc, aes(x=dbh_bin, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)

##Mature
mat_struc <- mat_struc %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
ggplot(mat_struc, aes(x=dbh_bin, y=n_stems, fill=spp)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)
ggplot(mat_struc, aes(x=dbh_bin, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_spp)
```

Compare the HDM infection in the mature component across the sites by looking at: basal area and stem count distributions parsed out by presence/absence of hdm infection and dmr. Generally trying to assess whether the mature components are similar infection sources. 
- Some spread in the proportion of the Hw basal area infected (37% at mk_1 to 100% at cr_1) but 7/8 sites>70 and 6 out of 8 > 80%.
- Mature components have different basal areas so its important to look at the magnitude (not just proportion) infected
- Graph comparing dmr across diameter classes is super useful. Sites with old mature components with severely infected big mature trees have most "inoculum load". 
```{r}
#Filter to just Hw trees
mat_hw <- mature_comp %>% filter(spp=="Hw") #mature trees

#Start simple. Calculate: basal area of infected trees at each site
mat_inf_ba_m2 <- mat_hw %>% filter(hdm_pa =="Y") %>% group_by(id_bec_age_f) %>% summarise(mat_inf_ba_m2 = sum(ba_m2))
site <- left_join(site, mat_inf_ba_m2, by="id_bec_age_f") #join variables back to site data
site <- site %>% mutate(p_mat_hw_ba_inf = mat_inf_ba_m2/mat_hw_ba_m2) #calculate the proportion of Hw basal area that's infected at each site

#Create a summary table. 
site %>% select(id_bec_age_f, mat_hw_ba_m2, mat_inf_ba_m2, p_mat_hw_ba_inf) %>% #summary table sorted by proportion of BA infected
  arrange(p_mat_hw_ba_inf) 

#Now break down basal area and # stems by site, presence/absece of hdm and diameter class
mat_hdmpa <- mat_hw %>% group_by(id_bec_age_f, hdm_pa, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% complete(dbh_bin, fill=list(ba_m2=0, n_stems=0))

##Graph this across diameter classes
mat_hdmpa <- mat_hdmpa %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
ggplot(mat_hdmpa, aes(x=dbh_bin, y=n_stems, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
ggplot(mat_hdmpa, aes(x=dbh_bin, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)

#Do the same but by DMR
mat_dmr <- mat_hw %>% group_by(id_bec_age_f, dmr_f2, dbh_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% complete(dbh_bin, fill=list(ba_m2=0, n_stems=0))

#Graph this across diameter classes
mat_dmr <- mat_dmr %>% filter(ba_m2>0) #Remove rows with 0 because R will plot all dbh intervals (up to 180cm)
ggplot(mat_dmr, aes(x=dbh_bin, y=n_stems, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
ggplot(mat_dmr, aes(x=dbh_bin, y=ba_m2, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
```

Compare infection levels and distribution of infection in the regen compoenent. 
- Note variable length transects were used and we need to account for this when comparing between sites. Calculating the proportion of the total basal area that is infected at a site isn't useful. i.e. 50% infection at site 1 vs 40% infection at site 2 isn't a meaningful comparison if site 1 and site 2 have different length transects. Deal with this in two ways: compare total magnitude (e.g. basal area infected) --> sites with longer transects are places where more spread has occured amd will likely have higher total magnitudes OR compare discrete intervals (e.g. basal area infected within 15m)
- There portion of Hw basal within 15m from infected wtih hdm ranges from ~0 - 72% but this spread is mostly caused by an outlier (ph_1) and the young site (mk_3). Between the other six sites, the range is 60-72%. 
- Notable from plots of ba ~ dist_y, coloured by dmr is that most infection is light. 
  - FOLLOW UP: some of trees at mi_1 that are severely infected are trees we thought might have been residuals. Check these and their cores?
```{r}
#Filter full dataset (not trees_comp which filters transects to <15m)
reg_hw <- trees %>% filter(tree_type == "regen" & spp=="Hw") #Hemlock regen trees 

#Start simple again, just calculate basal area infected at each site
reg_inf_ba_m2 <- reg_hw %>% filter(hdm_pa =="Y") %>% group_by(id_bec_age_f) %>% summarise(reg_inf_ba_m2 = sum(ba_m2))
reg_15_inf_ba_m2 <- regen_15 %>% filter(hdm_pa =="Y") %>% group_by(id_bec_age_f) %>% summarise(reg_15_inf_ba_m2 = sum(ba_m2))
site <- left_join(site, reg_inf_ba_m2, by="id_bec_age_f") #join variables back to site data
site <- left_join(site, reg_15_inf_ba_m2, by="id_bec_age_f")
site <- site %>% mutate(p_reg_15_ba_inf = reg_15_inf_ba_m2/reg_tot_ba_m2)

#Create a summary table. 
site %>% select(id_bec_age_f, reg_hw_ba_m2, reg_inf_ba_m2, reg_15_inf_ba_m2, p_reg_15_ba_inf) %>% 
  arrange(reg_15_inf_ba_m2) #summary table sorted by portion of BA infected within 15m of edge

#Create a plot relating basal area infected in mature component to basal area infected in the regen
ggplot(site, aes(x=mat_inf_ba_m2, y=reg_inf_ba_m2, color=id_bec_age_f)) + 
  geom_point()
## Can't really see anything, too few datapoints, but my sense is that inocukum pressure is going to be the key variable

#Do some plotting looking at HDM infection vs distance from the edge
##Break up dist_y_h into intervals. Need these to make barcharts
summary(reg_hw$dist_y_h)
reg_hw <- reg_hw %>% mutate(dist_y_h_bin = cut(dist_y_h, breaks = c(0, 5, 10, 15, 20, 25, 30, 35), include.lowest = TRUE))
levels(reg_hw$dist_y_h_bin)

##Summarize basal area and the # stems by site, presence/absence of hdm and distance from the edge (in the 5m intervals created above)
reg_hdmpa <- reg_hw %>% group_by(id_bec_age_f, hdm_pa, crown_class, dist_y_h_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% 
  complete(dist_y_h_bin, fill=list(ba_m2=0, n_stems=0))

##Plot nstems and basal area vs. distance from the edge, coloured by presence/absence of hdm infection
ggplot(reg_hdmpa, aes(x=dist_y_h_bin, y=n_stems, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_hdm_pa)
ggplot(reg_hdmpa, aes(x=dist_y_h_bin, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_hdm_pa)

##Do the same but with DMR. 
reg_dmr <- reg_hw %>% group_by(id_bec_age_f, dmr_f2, crown_class, dist_y_h_bin) %>% 
  summarise(ba_m2 = sum(ba_m2), n_stems = n()) %>% 
  complete(dist_y_h_bin, fill=list(ba_m2=0, n_stems=0))

##Plot nstems and basal area across distance from the edge, coloured by presence/absence of hdm infection
ggplot(reg_dmr, aes(x=dist_y_h_bin, y=n_stems, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)
ggplot(reg_dmr, aes(x=dist_y_h_bin, y=ba_m2, fill=dmr_f2)) + geom_bar(position="stack", stat="identity") + 
  facet_wrap(vars(id_bec_age_f)) + scale_fill_manual(values = colors_dmr_f2)

#Do some plotting to see if you can pick up patterns that relate the basal area of infected regen to site level variables
##Relating basal area infected by distance from the edge to site age
reg_inf_ag <- reg_hw %>% filter(hdm_pa=="Y") %>% 
  group_by(id_bec_age_f, age, dist_y_h_bin) %>% summarise(reg_inf_ba_m2 = sum(ba_m2)) %>% 
  complete(dist_y_h_bin, fill=list(reg_inf_ba_m2=0))
ggplot(reg_inf_ag, aes(x=dist_y_h_bin, y=reg_inf_ba_m2, color = age, group = id_bec_age_f)) + 
  geom_point() + geom_line()
##Relating basal area infected by distance from the edge to basal area infected in mature component
reg_inf_matinf <- reg_hw %>% filter(hdm_pa=="Y") %>% 
  group_by(id_bec_age_f, dist_y_h_bin) %>% summarise(reg_inf_ba_m2 = sum(ba_m2)) %>% 
  complete(dist_y_h_bin, fill=list(reg_inf_ba_m2=0))
reg_inf_matinf <- left_join(reg_inf_matinf, select(site, id_bec_age_f, mat_inf_ba_m2))
ggplot(reg_inf_matinf, aes(x=dist_y_h_bin, y=reg_inf_ba_m2, color = mat_inf_ba_m2, group = id_bec_age_f)) + 
  geom_point() + geom_line()
```



Extra code
```{r}
#Spread metrics
spread_metrics <- regen %>% filter(hdm_pa == "Y") %>% group_by(site_id, plot_id) %>% 
  summarise(q75 = quantile(dist_y_h, probs = c(0.75)), ba_inf=sum(ba_m2))
spread_metrics <- left_join(spread_metrics, select(site, age, site_id), by="site_id")

ggplot(spread_metrics, aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age")
ggplot(filter(spread_metrics, site_id!="mk_3"), aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age, minus outlier")


####################################

#Plot basal area by species at each site
ggplot(mature, aes(x=site_id, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity")
ggplot(regen, aes(x=site_id, y=ba_m2, fill=spp)) + geom_bar(position="stack", stat="identity")

#Plot basal area of infected vs. uninfected hemlock
mature_hw <- mature %>% filter(spp=="Hw")
regen_hw <- regen_15 %>% filter(spp=="Hw")

ggplot(mature_hw, aes(x=site_id, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity")  + 
  scale_fill_manual(values = c('#2c7bb6', '#fdae61', '#d7191c'))
ggplot(regen_hw, aes(x=site_id, y=ba_m2, fill=hdm_pa)) + geom_bar(position="stack", stat="identity") + 
  scale_fill_manual(values = c('#2c7bb6', '#fdae61', '#d7191c'))

#Start to build a dataframe at the transect level for plotting
tran <- trees %>% filter(tree_type=="regen") %>% select(site_id, plot_id) %>% 
  distinct() #step 1: two columns, site_id and plot_id
tran <- left_join(tran, select(site, age, cluster, bec_sz, site_id), 
                  by="site_id") #step 2: add site level variables
#step 3: calculate some metrics of spread
#metric 1: 75% percentiles of infected trees for each transect (i.e. the distance where 75% of the infected trees are between there and the mature component)
#metric 2: basal area infected
spread_metrics <- regen %>% filter(hdm_pa == "Y") %>% group_by(plot_id) %>% 
  summarise(q75 = quantile(dist_y_h, probs = c(0.75)), ba_inf=sum(ba_m2))
ba_15_inf <- regen %>% filter(hdm_pa == "Y" & dist_y_h<15) %>% group_by(plot_id) %>% 
  summarise(ba_15_inf = sum(ba_m2))
ba_15_tot <- regen %>% filter(dist_y_h<15) %>% group_by(plot_id) %>% summarise(ba_15_tot = sum(ba_m2))
tran <- left_join(tran, spread_metrics, by="plot_id")
tran <- left_join(tran, ba_15_inf, by="plot_id")
tran <- left_join(tran, ba_15_tot, by="plot_id")
tran <- tran %>% replace_na(list(q75 = 0, ba_inf = 0, ba_15_inf=0)) #replace NAs with 0
tran <- tran %>% mutate(p_ba_15_inf = ba_15_inf/ba_15_tot)

#Scatterplots
ggplot(tran, aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age")
ggplot(filter(tran, site_id!="mk_3"), aes(x=age, y=q75)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se = FALSE) + ggtitle("75 percentile dist of infected Hw vs age, minus outlier")
ggplot(tran, aes(x=age, y=ba_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age")
ggplot(filter(tran, site_id!="mk_3"), aes(x=age, y=ba_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age, minus outlier")
ggplot(tran, aes(x=age, y=p_ba_15_inf)) + geom_point(aes(col=site_id)) + 
  geom_smooth(method=lm, se=FALSE) + ggtitle("basal area infected Hw vs age, minus outlier")
```

