---
title: "Stem Mapped Figures"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
Read in data and load packages. 
"trees.geojson" is the output of the convert_distaz_to_points.R script. It contains all of the mature and regen trees as points and their associated data (e.g. DMR).  
"hdm_trimbpoints_2023.csv" are the trimble points from 2023 that have been cleaned up in GIS following the workflow in: /Users/hannosoutham/OneDrive - UBC (1)/Msc/GIS/ArcGIS_data_analysis_v1/hdm_arcgis_workflow.docx
"transect data_c.csv" is the cleaned transect data. Its used here to add the transect distances to the plot. 
```{r}
rm(list=ls(all=TRUE))

library(tidyverse)
library(sp)
library(sf)
library(tmap)
library(RColorBrewer)

trees <- read_sf("./data/workflow/trees_mapped.geojson")
trimb <- read_csv("./data/workflow/trimb_radjusted.csv")
tran <- read_csv('./data/cleaned/transect data_c.csv')
site <- read_csv('./data/cleaned/site data.csv')
```

Format the trees dataset. 
```{r}
class(trees)
summary(trees)

#check CRS, should be BC Albers NAD83 
st_crs(trees) #check CRS, should be BC Albers NAD83 = 3005

#Sort by site_id
trees <- trees %>% arrange(site_id)

# Convert to factor: spp, status, hdm_pa, b_lc, broom_pa, stem_pa, crown class, crown_cond, outside_10
trees <- trees %>% mutate(across(c(site_id, spp, status, hdm_pa, b_lc, broom_pa, 
                                   broom_pos, stem_pa, crown_class, crown_cond, outside_10, 
                                   assessed_by, tree_type, dmr_f), ~ as.factor(.)))

#Convert to integer: plot_id, dmrs
trees <- trees %>% mutate(across(
  c(plot_id, dmr_l, dmr_m, dmr_u, dmr), ~as.integer(.)))
str(trees) #check
```

Format the trimb dataset
```{r}
class(trimb)
summary(trimb)

#Convert to sf object
trimb <- st_as_sf(trimb, coords = c("plot_x_utm", "plot_y_utm"), crs=3005)
class(trimb)

#Check CRS is BC Albers NAD83 = 3005
st_crs(trimb)

#Sort by site_id
trimb <- trimb %>% arrange(site_id)

#Convert to pt_type and site_id to factor
trimb <- trimb %>% mutate(across(c(site_id, pt_type), ~ as.factor(.)))
str(trimb$pt_type)
str(trimb$site_id)

#Convert pt_id to a numeric
trimb <- trimb %>% mutate(pt_id = as.numeric(pt_id))
```

Format the site dataset
```{r}
#Sort by site_id
site <- site %>% arrange(site_id)

#Make factor variables factors: 
site <- site %>% mutate(across(c(site_id, cluster, bec_sz), ~ as.factor(.)))

#Calculate age (= years since harvest)
site <- site %>% mutate(age = (2023-yr_har))

#Create combined factor variable for plotting: site id + bec zone + age
site <- site %>% unite(col = bec_z_sz, sep = "", c("bec_z", "bec_sz"), remove = FALSE) %>% unite(col = id_bec_ag_f, sep = ", ", c("site_id", "bec_z_sz", "age"), remove = FALSE) %>% mutate(id_bec_ag_f = factor((id_bec_ag_f)))
```


Rotate the points at each site (in both trees and trimb) so they are all oriented with the mature component running vertically and the transects running horizontally. Code adapted with help from Chat GPT from: https://stackoverflow.com/questions/31873151/how-rotate-map-in-r
```{r}
#Calculate vector with how many degrees each site needs to be rotated. Want each transect to be az = 90deg. 
degrees <- trees %>% 
  mutate(rot_deg = case_when(tr_az<90 ~ (-1)*(tr_az-90), #define degree adjustment
                             tr_az>90 ~ (-1)*(tr_az-90))) %>% 
  filter(str_detect(tree_id, "r")) %>% #filter to just regen trees
  group_by(site_id) %>% 
  distinct(rot_deg) %>% # get unique values for each site
  pull(rot_deg) #make vector

#Define a list of centre coordinate corresponding to each site. Use the centroid of all the stem mapped trees.
center_coords <- trees %>% group_by(site_id) %>% 
  summarise(geometry = st_centroid(st_combine(geometry))) %>% st_geometry()

#Split datasets, one for each site. This results in list of sf objects, one for each site. For the point datasets (trees and trimb) all the points at a site have been combined to a single feature (corresponding to a sf class change of POINT to MULTIPOINT). 
pts_trees <- trees %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()
pts_trimb <- trimb %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()

#Define a function to rotate a spatial feature. x = an sf object, center_coords = the centre coordinates to anchor each site, degrees=the number of degrees to rotate the object
rotate_sf <- function(x, center_coords, degrees) {
  radians <- degrees * pi/180 #convert to radians
  transform_matrix <- matrix(c(cos(radians), sin(radians), #the transformation matrix
                                 -sin(radians), cos(radians)), 2, 2)
  rot_site <- ((x - center_coords) * transform_matrix) + center_coords #the actual math
  return(rot_site) #output
}

#Map function applies another function to each item in a vector successively. Here, it rotates each site level MULTIPOINT sf object by the number of degrees in the "degrees" vector. 
trees_rot <- Map(rotate_sf, pts_trees, center_coords, degrees)

trimb_rot <- Map(rotate_sf, pts_trimb, center_coords, degrees)

#This output is a new list of sf multipoint objects. Combine these back together into a dataframe of x, y coordinates. 
trees_rot <- do.call(rbind, trees_rot) #do.call applies function to each item in a list
trees_rot <- as.data.frame(trees_rot) %>% rename(X=V1, Y=V2)

trimb_rot <- do.call(rbind, trimb_rot) #do.call applies function to each item in a list
trimb_rot <- as.data.frame(trimb_rot) %>% rename(X=V1, Y=V2)

#Then associate back to the attribute data associated with each point
trees_att <- trees %>% st_drop_geometry()
trees_rot <- trees_att %>% mutate(X=trees_rot$X, Y=trees_rot$Y) 
trees_rot <- st_as_sf(trees_rot, coords = c("X", "Y"), crs=3005)

trimb_att <- trimb %>% st_drop_geometry()
trimb_rot <- trimb_att %>% mutate(X=trimb_rot$X, Y=trimb_rot$Y) 
trimb_rot <- st_as_sf(trimb_rot, coords = c("X", "Y"), crs=3005)
```

Make the stem mapped plot. 
```{r}
# Filter out mature trees that were mapped outside of 10m. Only happened at a few sites. 
levels(trees_rot$outside_10)
trees_rot <- trees_rot %>% filter(outside_10 == "N" | is.na(outside_10))

#Create the footprints of the each site:
tran_pts <- trimb_rot %>% #filter trimb_points to just transect starts and ends
  filter(pt_type %in% c("r adj tran start", "tran end"))
rot_points <- rbind(select(trees_rot, site_id), select(tran_pts, site_id)) #combine trees_rot and trimb_rot datasets
rot_points <- rot_points %>% group_by(site_id) %>% #split rotated points into separate datasets for each site
  summarise(geometry=st_combine(geometry)) %>% st_geometry()
fp_rot <- Map(st_bbox, rot_points) #create bounding box around each of these sets of points

# Calculate the maximum differences in x and y directions among all bounding boxes
# With help from ChatGPT
# Here, sapply applies the same function over each item in a list and outputs a vector. function defines what is going to be allied to each list item. Each item is called 'bbox' so it calls that item and then specifies which element e.g. bbox[1] = xmin. max() takes that vector and finds the biggest value. 
fp_rot[[1]][1] # just an example of how you access individual bbox elements. This accesses the first element (xmin) of the first bounding box
max_x_diff <- max(sapply(fp_rot, function(bbox) bbox[3] - bbox[1]))
max_y_diff <- max(sapply(fp_rot, function(bbox) bbox[4] - bbox[2]))
# Adjust each bounding box to have the maximum differences
# Here lapply applies a function to each item in a list. {} defines the function. In the function it finds the centre of xmin and xmax and then adds/subtracts half of what the extend should be (max_x_dff).
adjusted_fp_rot <- lapply(fp_rot, function(bbox) {
  adj_xmin <- ((bbox[1] + bbox[3])/2) - (max_x_diff/2) #calculate new xmin
  adj_ymin <- ((bbox[2] + bbox[4])/2) - (max_y_diff/2) #calculate new ymin
  bbox[1] <- adj_xmin #reassign xmin value
  bbox[2] <- adj_ymin #reassign ymin value
  bbox[3] <- adj_xmin + max_x_diff  # redefine xmax
  bbox[4] <- adj_ymin + max_y_diff  # redefine ymax
  
  bbox
})

# Print adjusted bounding boxes
max_x_diff; max_y_diff
adjusted_fp_rot
adjusted_fp_rot[[2]][2] - adjusted_fp_rot[[2]][4]

#Turn this back into an sf object that can be plotted
adjusted_fp_rot <- Map(st_as_sfc, adjusted_fp_rot) #turn bbox object into sfc object
adjusted_fp_rot <- Map(st_as_sf, adjusted_fp_rot) #trun sfc object into sf object
adjusted_fp_rot <- do.call(rbind, adjusted_fp_rot) #combine sf objects to single object
sites <- trees %>% st_drop_geometry() %>% distinct(site_id) #get column of site_ids
adjusted_fp_rot <- adjusted_fp_rot %>% mutate(site_id=sites$site_id) #add comlumn of site_ids to sf object
st_crs(adjusted_fp_rot) = 3005 #define crs

#Create the transect lines for the site
tran_pts <- trimb_rot %>% #filter to just transect starts and ends
  filter(pt_type %in% c("r adj tran start", "tran end"))
tran_line_att <- tran_pts %>% st_drop_geometry() #save attributes
tran_lines <- tran_pts %>% group_by(pt_id) %>% # draw lines between start and end. Creates two lines one from start to end and one from end to start.
  summarize() %>% st_cast("LINESTRING")
tran_lines <- st_buffer(tran_lines, dist = 2.5, endCapStyle = 'FLAT') # buffer
tran_lines <- left_join(tran_lines, tran_line_att, by="pt_id")
tran_lines <- tran_lines %>% filter(pt_type=="tran end")
tran_lines <- left_join(tran_lines, select(tran, transect_id, tr_leng), by=c("pt_id"="transect_id"))

#Create horizontal lines marking transect distances at 5m intervals
#Up to 6 lines/transect, 2pts/line, 10 sites, 3 transects/site
# #pts = 3*8*6*2 = 288 pts
max(tran$tr_leng) #find longest transect length. 32.8m (cr_1)
# Create new fake points
dist_x <- rep(c(-2.5, 2.5), 3*length(site$site_id)*6) #define vector of x coords
dist_y <- rep(c(5, 5, 10, 10, 15, 15, 20, 20, 25, 25, 30, 30), 3*length(site$site_id)) #define vector of y coordinates
dist_y <- sort(dist_y)
ids <- sample(1:500, length(site$site_id)*3*6, replace = FALSE) #generate 80 random unique ids
ids <- sort(rep(ids, 2))
transect_id <- pull(tran, transect_id)
transect_id <- rep(transect_id, 2)
transect_id <- sort(transect_id)
transect_id <- rep(transect_id, 6)
dist_mark <- tibble(ids, transect_id, dist_x, dist_y)

#calculate distance and degree adjustment from transect line
dist_mark <- dist_mark %>% mutate(dist = (dist_x^2 + dist_y^2)^0.5, az_adj = atan(dist_x/dist_y)*(180/pi))

#Because we are working with points, we know all transects are oriented to 90degrees. Azimuth to each point
dist_mark <- dist_mark %>% mutate(az = (90 + az_adj)%%360)

#Add site_id and transect length to this dataframe
dist_mark <- left_join(dist_mark, select(tran, transect_id, site_id, tr_leng), by="transect_id")

#Add column to check if dist_y > transect length. If it is, then drop the row
dist_mark <- dist_mark %>% mutate(diff_mark_tr = tr_leng-dist_y) %>% filter(diff_mark_tr > 0)

#Join these points to transect starts
ts_pts <- trimb_rot %>% filter(pt_type == "r adj tran start")
ts_coords <- st_coordinates(ts_pts) %>% as.data.frame()
ts_coords <- mutate(ts_coords, transect_id = ts_pts$pt_id)
dist_mark <- left_join(dist_mark, ts_coords, by="transect_id")

#Define function from Liam to generate points from from dist and az readings
polar_to_XY <- function(azimuth,
                        distance,
                        xcenter,
                        ycenter,
                        shape_file = TRUE,
                        crs) {
  
  if((max(azimuth) - min(azimuth) < 2*pi) == TRUE){
    print("WARNING: This function assumes azimuth is in degrees, please check")
  }
  angle = azimuth * pi/180
  #Convert to radians
  #angle = angle*pi/180
  angle = 2*pi - (angle - pi/2)
  x = xcenter + distance * cos(angle)
  y = ycenter + distance * sin(angle)
  
  #define output point locations
  tree_locations <- data.frame(X = x, Y = y)
  
  if(shape_file == T){# output a shape file of the tree locations
    print(paste("creating spatial points layer from tree locations. CRS is:", crs))
    
    tree_locations <- sf::st_as_sf(tree_locations, coords = c("X", 'Y'), crs = crs)
    
  }
  else{
    tree_locations
  }
  return(tree_locations)
}

#Run function
crs <- 3005
dist_mark <- dist_mark %>%
  mutate(polar_to_XY(azimuth = az, distance = dist,
                     xcenter = X, ycenter = Y, crs = crs,
                     shape_file = TRUE))

#Convert it to a spatial object:
dist_mark <- st_as_sf(dist_mark)
class(dist_mark)

#Create lines from each set of two points that mark intervals on the transect
dist_mark_att <- dist_mark %>% st_drop_geometry() #save attributes
dist_mark <- dist_mark %>% group_by(ids) %>% # draw lines between two points. Creates two lines one from start to end and one from end to start.
  summarize() %>% st_cast("LINESTRING")
dist_mark <- left_join(dist_mark_att, dist_mark, by="ids")
dist_mark <- st_as_sf(dist_mark)

#Create some edge lines:
levels(trimb_rot$pt_type)
edge_lines <- trimb_rot %>% #filter to points that define edge ("edge" and "r adj tran start")
  filter(pt_type %in% c("edge", "r adj tran start"))
edge_line_coords <- st_coordinates(edge_lines) %>%  #extract coordinates
  as.data.frame() %>% mutate(diff=(X-Y))
edge_lines <- edge_lines %>% mutate(diff=edge_line_coords$diff)
edge_lines <- edge_lines %>% group_by(site_id) %>% arrange(diff) %>% # draw lines between start and end
  summarise(do_union = FALSE) %>% st_cast("LINESTRING")

#Add a layer to visualize seed load
#sl <- trees_rot %>% filter(tree_type=="regen" & sl>0)

#Add site data to each of the objects being plotted
adjusted_fp_rot <- left_join(adjusted_fp_rot, select(site, site_id, id_bec_ag_f), by ="site_id")
edge_lines <- left_join(edge_lines, select(site, site_id, id_bec_ag_f), by ="site_id")
tran_lines <- left_join(tran_lines, select(site, site_id, id_bec_ag_f), by ="site_id")
tran_pts <- left_join(tran_pts, select(site, site_id, id_bec_ag_f), by ="site_id")
dist_mark <- left_join(dist_mark, select(site, site_id, id_bec_ag_f), by ="site_id")
trees_rot <- left_join(trees_rot, select(site, site_id, id_bec_ag_f), by ="site_id")
#sl <- left_join(sl, select(site, site_id, id_bec_ag_f), by ="site_id")

#Make new version of dmr_f for graphing that simplifies some of the ratings
trees_rot <- trees_rot %>% mutate(dmr_f_graph = case_when(dmr_f == "-" ~ "non Hw", dmr_f %in% c("1", "2") ~ "1-2", dmr_f %in% c("3", "4") ~ "3-4", dmr_f %in% c("5", "6") ~ "5-6", .default = dmr_f)) %>% mutate(dmr_f_graph = factor(dmr_f_graph, levels = c("non Hw", "DU", "0", "DI", "IBLC", "1-2", "3-4", "5-6")))

#Define colour pallette
# https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3
levels(trees_rot$dmr_f_graph)
dmr_colour <- c('#cccccc','#4575b4', '#74add1', '#c994c7','#fdbb84','#ef6548','#d7301f','#990000')
sl_colour <- brewer.pal(n=9, name="Greys")
brewer.pal.info

# Plot it all together
# pdf("./figures/trees_25Jul2024.pdf", width=11, height=8.5)
tmap_mode("plot")
tm_shape(adjusted_fp_rot, is.master = TRUE) + #define map extent using site footprints
  tm_facets(by = "id_bec_ag_f", free.scales = FALSE, free.coords = TRUE, ncol = 4) +
  tm_polygons(col="white", border.col = NULL) +
tm_shape(edge_lines) + tm_lines(col="blue", lty=2) + #add edge lines
tm_shape(tran_lines) + tm_polygons(alpha=0.25) + 
  tm_text("tr_leng", ymod = 1.7) + #add transects
tm_shape(tran_pts) + tm_symbols(col="blue", shape=3, size=0.1) + #transect start and end points
tm_shape(dist_mark) + tm_lines(lty = 3, alpha = 0.7) +
tm_shape(trees_rot) + tm_symbols(size = "dbh", scale = 2, col="dmr_f_graph", palette = dmr_colour, border.lwd = NA, title.col = "HDM status", legend.size.is.portrait = TRUE, legend.col.is.portrait = FALSE) + #add tree points
tm_legend(legend.outside=TRUE, legend.outside.position = "bottom", legend.position=c(0.4, 0.2), 
                                 legend.stack="horizontal")
dev.off()

#Seed load addition
#tm_shape(sl) + tm_symbols(size = 2, col="sl", border.lwd = NA, title.col = "seed load", legend.size.is.portrait = TRUE, legend.col.is.portrait = FALSE)
```

