---
title: "Stem Mapped Figures"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
##############
READ ME
This script generates stem maps of each HDM site. The core starting objects are
the output of the 'seed load.Rmd' script (trees_sl.csv) and processed version 
of the trimble points (trimb_radjusted.csv). 

The {sf} and {sp} packages are used to process the spatial data and the {tmap}
package is used to make the actual plots. {tmap} has a similar syntax to ggplot and a similar logic to ArcGIS. You can add layers and slowly build up a map. 

The script has two major sections.The first section defines a function to 
rotate the points at each site a fixed number of degrees so they all end up 
pointing the same direction in the stem maps. The second section does the 
actual plotting. 
##############x

#Setup
```{r}
rm(list=ls(all=TRUE))

library(tidyverse)
library(here)
library(sp)
library(sf)
library(tmap)
library(RColorBrewer)

trees <- read_csv("./data/workflow/trees_sl.csv")
trimb <- read_csv("./data/workflow/trimb_radjusted.csv")
tran <- read_csv('./data/cleaned/transect data_c.csv')
site <- read_csv('./data/cleaned/site data.csv')
```

Format the trees dataset. 
```{r}
#Convert to a sf, point feature object
trees <- st_as_sf(trees, coords = c("X", "Y"), crs = 3005)

class(trees)
summary(trees)

#Check CRS, should be BC Albers NAD83 
st_crs(trees) #check CRS, should be BC Albers NAD83 = 3005

#Sort by site_id
trees <- trees %>% arrange(site_id)

#Convert to factor: spp, status, hdm_pa, b_lc, broom_pa, stem_pa, crown class, 
#crown_cond, outside_10
trees <- trees %>% 
  mutate(across(c(site_id, spp, status, hdm_pa, b_lc, broom_pa, broom_pos, 
                  stem_pa, crown_class, crown_cond, outside_10, 
                  assessed_by, tree_type, tree_type_2, dmr_f,
                  sim_tree), ~ as.factor(.)))

#Convert to integer: plot_id, dmrs
trees <- trees %>% mutate(across(
  c(plot_id, dmr_l, dmr_m, dmr_u, dmr), ~as.integer(.)))
str(trees) #check
```

Format the trimb dataset
```{r}
class(trimb)
summary(trimb)

#Convert to sf object
trimb <- st_as_sf(trimb, coords = c("plot_x_utm", "plot_y_utm"), crs=3005)
class(trimb)

#Check CRS is BC Albers NAD83 = 3005
st_crs(trimb)

#Sort by site_id
trimb <- trimb %>% arrange(site_id)

#Convert to pt_type and site_id to factor
trimb <- trimb %>% mutate(across(c(site_id, pt_type), ~ as.factor(.)))
str(trimb$pt_type)
str(trimb$site_id)

#Convert pt_id to a numeric
trimb <- trimb %>% mutate(pt_id = as.numeric(pt_id))
```

Format the site dataset
```{r}
#Sort by site_id
site <- site %>% arrange(site_id)

#Make factor variables factors: 
site <- site %>% mutate(across(c(site_id, cluster, bec_sz), ~ as.factor(.)))

#Calculate age (= years since harvest)
site <- site %>% mutate(age = (2023-yr_har))

#Create combined factor variable for plotting: site id + bec zone + age
site <- site %>% 
  unite(col = bec_z_sz, sep = "", c("bec_z", "bec_sz"), remove = FALSE) %>% 
  unite(col = id_bec_ag_f, sep = ", ", 
        c("site_id", "bec_z_sz", "age"), remove = FALSE) %>% 
  mutate(id_bec_ag_f = factor((id_bec_ag_f)))
```

#Rotate points
Rotate the points at each site (in both trees and trimb) so they are all 
oriented with the mature component running vertically and the transects running 
horizontally. Code adapted with help from Chat GPT and from: https://stackoverflow.com/questions/31873151/how-rotate-map-in-r
```{r}
#Calculate vector with how many degrees each site needs to be rotated. Want 
#each transect to be az = 90deg. 
degrees <- trees %>% 
  #define degree adjustment
  mutate(rot_deg = case_when(tr_az<90 ~ (-1)*(tr_az-90), 
                             tr_az>90 ~ (-1)*(tr_az-90))) %>% 
  #filter to just regen trees
  filter(str_detect(tree_id, "r")) %>%
  #get unique values for each site
  group_by(site_id) %>% 
  distinct(rot_deg) %>% 
  #pull out that vector of adjusment values for each site
  pull(rot_deg)

#Define a list of centre coordinate corresponding to each site. Use the 
#centroid of all the stem mapped trees.
center_coords <- trees %>% group_by(site_id) %>% 
  summarise(geometry = st_centroid(st_combine(geometry))) %>% st_geometry()

#Split datasets, one for each site. This results in list of sf objects, one for 
#each site. For the point datasets (trees and trimb) all the points at a site 
#have been combined to a single feature (corresponding to a sf class change of 
#POINT to MULTIPOINT). 
pts_trees <- trees %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()
pts_trimb <- trimb %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()

#Define a function to rotate a spatial feature. x = an sf object, 
#center_coords = the centre coordinates to anchor each site, degrees=the number 
#of degrees to rotate the object
rotate_sf <- function(x, center_coords, degrees) {
  #convert to radians
  radians <- degrees * pi/180 
  #the transformation matrix
  transform_matrix <- matrix(c(cos(radians), sin(radians), 
                                 -sin(radians), cos(radians)), 2, 2)
  #the actual math
  rot_site <- ((x - center_coords) * transform_matrix) + center_coords 
  return(rot_site) #output
}

#Map function applies another function to each item in a vector successively. 
#Here, it rotates each site level MULTIPOINT sf object by the number of degrees 
#in the "degrees" vector. 
trees_rot <- Map(rotate_sf, pts_trees, center_coords, degrees)

trimb_rot <- Map(rotate_sf, pts_trimb, center_coords, degrees)

#This output is a new list of sf multipoint objects. Combine these back 
#together into a dataframe of x, y coordinates. 
#do.call applies function to each item in a list
trees_rot <- do.call(rbind, trees_rot) 
trees_rot <- as.data.frame(trees_rot) %>% rename(X=V1, Y=V2)

trimb_rot <- do.call(rbind, trimb_rot) 
trimb_rot <- as.data.frame(trimb_rot) %>% rename(X=V1, Y=V2)

#Then associate back to the attribute data associated with each point
trees_att <- trees %>% st_drop_geometry()
trees_rot <- trees_att %>% mutate(X=trees_rot$X, Y=trees_rot$Y) 
trees_rot <- st_as_sf(trees_rot, coords = c("X", "Y"), crs=3005)

trimb_att <- trimb %>% st_drop_geometry()
trimb_rot <- trimb_att %>% mutate(X=trimb_rot$X, Y=trimb_rot$Y) 
trimb_rot <- st_as_sf(trimb_rot, coords = c("X", "Y"), crs=3005)
```

#Make the stem maps
##Layer 1: tree points
```{r}
#Filter out mature trees that were mapped outside of 10m. Only happened 
#at a few sites. 
trees_rot <- trees_rot %>% filter(outside_10 == "N" | is.na(outside_10))
```

##Layer 2: trimble points
```{r}
#Filter to trimble points identified as for graphing
trimb_rot <- trimb_rot %>% filter(graph == "Y")

#Take a look at the levels of pt_type, used frequently below
levels(trimb_rot$pt_type)
```

##Layer 3: Site footprints
```{r}
##These are standardized to be the same size across all sites to ensure the 
##scale is the the same. 

#Create the footprints of the each site:
##Split rotated points into separate datasets for each site
fp_points <- trees_rot %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()

##Create bounding box around each of these sets of points
fp_rot <- Map(st_bbox, fp_points)

#Calculate the maximum differences in x and y directions among all bounding 
#boxes
#With help from ChatGPT
#Here, sapply applies the same function over each item in a list and outputs 
#a vector. 
#Function defines what is going to be applied to each list item. Each item is 
#called 'bbox' so it calls that item and then specifies which element 
#e.g. bbox[1] = xmin. max() takes that vector and finds the biggest value. 

#Example of how you access individual bbox elements. This accesses the first 
#element (xmin) of the first bounding box
fp_rot[[1]][1] 

max_x_diff <- max(sapply(fp_rot, function(bbox) bbox[3] - bbox[1]))
max_y_diff <- max(sapply(fp_rot, function(bbox) bbox[4] - bbox[2]))

#Adjust each bounding box to have the maximum differences
#Here lapply applies a function to each item in a list. {} defines the 
#function. In the function it finds the centre of xmin and xmax and then 
#adds/subtracts half of what the extend should be (max_x_dff).
adjusted_fp_rot <- lapply(fp_rot, function(bbox) {
  adj_xmin <- ((bbox[1] + bbox[3])/2) - (max_x_diff/2) #calculate new xmin
  adj_ymin <- ((bbox[2] + bbox[4])/2) - (max_y_diff/2) #calculate new ymin
  bbox[1] <- adj_xmin #reassign xmin value
  bbox[2] <- adj_ymin #reassign ymin value
  bbox[3] <- adj_xmin + max_x_diff  # redefine xmax
  bbox[4] <- adj_ymin + max_y_diff  # redefine ymax
  
  bbox
})

#Print adjusted bounding boxes
max_x_diff; max_y_diff
adjusted_fp_rot
adjusted_fp_rot[[2]][2] - adjusted_fp_rot[[2]][4]

#Turn this back into an sf object that can be plotted
##Rurn bbox object into sfc object
adjusted_fp_rot <- Map(st_as_sfc, adjusted_fp_rot) 

##Turn sfc object into sf object
adjusted_fp_rot <- Map(st_as_sf, adjusted_fp_rot) 

##Combine sf objects to single object
adjusted_fp_rot <- do.call(rbind, adjusted_fp_rot) 

##Get column of site_ids
sites <- trees %>% st_drop_geometry() %>% distinct(site_id) 

##Add comlumn of site_ids to sf object
adjusted_fp_rot <- adjusted_fp_rot %>% mutate(site_id=sites$site_id) 
st_crs(adjusted_fp_rot) = 3005 #define crs
```

##Layers 4, 5, 6: Transects
Transect start and end points
```{r}
#Pull out transect start and end points separately
#Will use these objects for plotting below
te_pts <- trimb_rot %>% 
  filter(pt_type == "tran start" & 
           graph == "Y")
ts_pts <- trimb_rot %>% 
  filter(pt_type == "tran end" & 
           graph == "Y")

#Add transect length to transect end points
x <- tran %>% select(transect_id, tr_leng)
te_pts <- left_join(te_pts, x, by = join_by(pt_id == transect_id))
```

Transect footprints
```{r}
#Generate transect footprints from start and end points
##Pull out transect start and end points together
##Will use this object to generate footprints
tran_pts <- trimb_rot %>% 
  filter(pt_type %in% c("tran start", "tran end"))

##Save attributes.  Use transect start attributes to represent footprints
tran_line_att <- tran_pts %>% st_drop_geometry() %>% 
  filter(pt_type == "tran start")

##Draw lines between start and end. 
##group_by() and summarize() ensure only one line is created for each pair
tran_lines <- tran_pts %>% group_by(pt_id) %>% 
  summarize() %>% st_cast("LINESTRING")

##Buffer
tran_lines <- st_buffer(tran_lines, dist = 2.5, endCapStyle = 'FLAT')

##Join back to attributes.
tran_lines <- left_join(tran_lines, tran_line_att, by="pt_id")

##Add transect length from transect data
tran_lines <- left_join(tran_lines, select(tran, transect_id, tr_leng), 
                        by=c("pt_id"="transect_id"))
```

Transect distance markers
```{r}
#Create horizontal lines marking transect distances at 5m intervals
##Up to 10 lines/transect, 2pts/line, 11 sites, 3 transects/site

##Find the longest transect length. 51m (cr_3)
max(tran$tr_leng)

##Calculate how many 5m intervals are needed for the max length transect
n_5_int <- round(max(tran$tr_leng)/5, 0)

##Create new fake points
###Define vector of x coords
dist_x <- rep(c(-2.5, 2.5), 3*length(site$site_id)*n_5_int)

###Define vector of y coordinates
dist_y <- rep(c(5, 5, 10, 10, 15, 15, 20, 20, 25, 25, 30, 30, 35, 35, 40, 40, 
                45, 45, 50, 50), 
              3*length(site$site_id)) 
dist_y <- sort(dist_y)

###Generate random unique ids
ids <- sample(1:500, length(site$site_id)*3*n_5_int, replace = FALSE) 
ids <- sort(rep(ids, 2))
transect_id <- pull(tran, transect_id)
transect_id <- rep(transect_id, 2)
transect_id <- sort(transect_id)
transect_id <- rep(transect_id, n_5_int)
dist_mark <- tibble(ids, transect_id, dist_x, dist_y)

##Calculate distance and degree adjustment from transect line
dist_mark <- dist_mark %>% mutate(dist = (dist_x^2 + dist_y^2)^0.5, 
                                  az_adj = atan(dist_x/dist_y)*(180/pi))

##Calculate angle from transect start point to the distance marker
dist_mark <- dist_mark %>% mutate(az = (90 + az_adj)%%360)

##Add site_id and transect length to this dataframe
dist_mark <- left_join(dist_mark, 
                       select(tran, transect_id, site_id, tr_leng), 
                       by="transect_id")

##Add column to check if dist_y > transect length. If it is, then drop the row
dist_mark <- dist_mark %>% 
  mutate(diff_mark_tr = tr_leng-dist_y) %>% 
  filter(diff_mark_tr > 0)

##Join these points to transect starts used for stem mapping
ts_pts_sm <- trimb_rot %>% 
  filter(pt_type == "tran start" & stem_map == "Y")
ts_coords <- st_coordinates(ts_pts_sm) %>% as.data.frame()
ts_coords <- mutate(ts_coords, transect_id = ts_pts$pt_id)
dist_mark <- left_join(dist_mark, ts_coords, by="transect_id")

##Define function from Liam to generate points from from dist and az readings
polar_to_XY <- function(azimuth,
                        distance,
                        xcenter,
                        ycenter,
                        shape_file = TRUE,
                        crs) {
  
  if((max(azimuth) - min(azimuth) < 2*pi) == TRUE){
    print("WARNING: This function assumes azimuth is in degrees, please check")
  }
  angle = azimuth * pi/180
  #Convert to radians
  #angle = angle*pi/180
  angle = 2*pi - (angle - pi/2)
  x = xcenter + distance * cos(angle)
  y = ycenter + distance * sin(angle)
  
  #define output point locations
  tree_locations <- data.frame(X = x, Y = y)
  
  if(shape_file == T){# output a shape file of the tree locations
    print(paste("creating spatial points layer from tree locations. CRS is:", crs))
    
    tree_locations <- sf::st_as_sf(tree_locations, coords = c("X", 'Y'), crs = crs)
    
  }
  else{
    tree_locations
  }
  return(tree_locations)
}

##Run function
crs <- 3005
dist_mark <- dist_mark %>%
  mutate(polar_to_XY(azimuth = az, distance = dist,
                     xcenter = X, ycenter = Y, crs = crs,
                     shape_file = TRUE))

##Convert it to a spatial object:
dist_mark <- st_as_sf(dist_mark)
class(dist_mark)

##Create lines from each set of two points that mark intervals on the transect
###Save attributes
dist_mark_att <- dist_mark %>% st_drop_geometry() 

###Draw lines between two points. Creates two lines one from start to end and
###one from end to start.
dist_mark <- dist_mark %>% group_by(ids) %>% 
  summarize() %>% st_cast("LINESTRING")
dist_mark <- left_join(dist_mark_att, dist_mark, by="ids")
dist_mark <- st_as_sf(dist_mark)
```

##Layer 7: Edge lines
```{r}
#Create some edge lines:
levels(trimb_rot$pt_type)

##Filter to points that define edge ("edge" and "r adj tran start")
edge_lines <- trimb_rot %>% 
  filter(pt_type %in% c("edge", "tran start"))

##Extract coordinates
edge_line_coords <- st_coordinates(edge_lines) %>%  
  as.data.frame() %>% mutate(diff=(X-Y))

##Draw lines between start and end
edge_lines <- edge_lines %>% mutate(diff=edge_line_coords$diff)
edge_lines <- edge_lines %>% group_by(site_id) %>% arrange(diff) %>% 
  summarise(do_union = FALSE) %>% st_cast("LINESTRING")
```

##Plotting prep
Add site id and factor with site_id, bec and age to each of the objects and 
make a new version of DMR for plotting.
```{r}
#Add site id, bec and age
adjusted_fp_rot <- left_join(adjusted_fp_rot, 
                             select(site, site_id, id_bec_ag_f), by ="site_id")
edge_lines <- left_join(edge_lines, 
                        select(site, site_id, id_bec_ag_f), by ="site_id")
tran_lines <- left_join(tran_lines, 
                        select(site, site_id, id_bec_ag_f), by ="site_id")
ts_pts <- left_join(ts_pts, 
                      select(site, site_id, id_bec_ag_f), by ="site_id")
te_pts <- left_join(te_pts, 
                      select(site, site_id, id_bec_ag_f), by ="site_id")
dist_mark <- left_join(dist_mark, 
                       select(site, site_id, id_bec_ag_f), by ="site_id")
trees_rot <- left_join(trees_rot, 
                       select(site, site_id, id_bec_ag_f), by ="site_id")

#Make new version of dmr_f for graphing that simplifies some of the ratings
levels(trees_rot$dmr_f)
trees_rot <- trees_rot %>% 
  mutate(dmr_f_graph = case_when(dmr_f == "-" ~ "non-hemlock", 
                                 dmr_f %in% c("1", "2") ~ "1-2", 
                                 dmr_f %in% c("3", "4") ~ "3-4", 
                                 dmr_f %in% c("5", "6") ~ "5-6",
                                 .default = dmr_f)) %>% 
  mutate(dmr_f_graph = factor(dmr_f_graph, 
                              levels = c("non-hemlock", "DU", "0", 
                                         "IBLC", "1-2", 
                                         "3-4", "5-6", "DI")))
```

Define a new factor that describes whether a tree was simulated or not to use
as the shape aesthetic in these plots. 
```{r}
trees_rot <- trees_rot %>% 
  mutate(sim_tree_g = if_else(sim_tree == "Y", "simulated", "measured")) %>% 
  mutate(sim_tree_g = factor(sim_tree_g, levels = c("measured", "simulated")))
levels(trees_rot$sim_tree_g)
```

Define shapes and colours
```{r}
#DMR colour palette
# https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3
levels(trees_rot$dmr_f_graph)
color_dmr <- c("black", "#1F78B4", "#A6CEE3", 
                       brewer.pal(4, "Oranges"), "hotpink")

#Data source
levels(trees_rot$sim_tree_g)
shape_sim <- c(16, 15)

#Available colour palettes in tmap:
# tmaptools::palette_explorer()
```

##Plot 1: standard stem map
```{r}
p <- tmap_mode("plot") +
  #Define map extent using site footprints
  tm_shape(adjusted_fp_rot, is.master = TRUE) +
    tm_facets(by = "site_id", free.scales = FALSE, 
              free.coords = TRUE, ncol = 2) +
    tm_polygons(col="white", border.col = NULL) +
  #Edge lines
  tm_shape(edge_lines) + 
    tm_lines(col="blue", lty=2) +
  #Transect start/end points, footprints and distance markers
  tm_shape(tran_lines) + 
    tm_polygons(alpha=0.25) + 
  #Transect start and end points
  tm_shape(ts_pts) + 
    tm_symbols(col="blue", shape=3, size=0.1) + 
  tm_shape(te_pts) + 
    tm_symbols(col="blue", shape=3, size=0.1) +
  tm_shape(dist_mark) + 
    tm_lines(lty = 3, alpha = 0.7) +
  #Tree points
  tm_shape(trees_rot) + 
    tm_symbols(col ="dmr_f_graph",
               palette = color_dmr,
               shape =  "sim_tree_g",
               shapes =  shape_sim, 
               size = "dbh", scale = 2,
               title.col = "Adapted DMR",
               legend.col.is.portrait = T,
               title.size = "DBH (cm)",
               legend.size.is.portrait = T,
               title.shape = "Tree type",
               legend.shape.is.portrait = T) +
  tm_layout(inner.margins = 0.01, panel.label.size = 2,
            panel.label.bg.color = "white",
            legend.outside=T, 
            legend.outside.position = "right",
            legend.position=c(0.1, 0.3), 
            legend.stack="vertical",
            legend.title.size = 2,
            legend.text.size = 1.5,
            legend.bg.color = "white")

# pdf("./figures/trees_13Nov2024.pdf", width=8.5, height=11)
# p 
# dev.off()
```

##Plot 2: stem map with seed load overlay
Create a new object that is just points of trees where seed load is defined
```{r}
trees_sl <- trees_rot %>% 
  filter(!is.na(sl_i_tot_rm))

#These are all live Hw trees, about half measured and half simulated
summary(trees_sl$tree_type_2)
summary(trees_sl$status)
summary(trees_sl$spp)
```

Break up objects into two sets. One for a figure in the main text and one for 
one to go in the supplementary material. 
```{r}
sites1 <- c("cr_1", "cr_3", "mi_1", "mk_2", "ph_1", "ph_2")
sites2 <- c("cr_2", "mi_2", "mk_1", "mk_3", "ph_3")

#Footprints
adjusted_fp_rot1 <- adjusted_fp_rot %>% 
  filter(site_id %in% sites1)
adjusted_fp_rot2 <- adjusted_fp_rot %>% 
  filter(site_id %in% sites2)

#Edge lines
edge_lines1 <- edge_lines %>% 
  filter(site_id %in% sites1)
edge_lines2 <- edge_lines %>% 
  filter(site_id %in% sites2)

#Transect footprints
tran_lines1 <- tran_lines %>% 
  filter(site_id %in% sites1)
tran_lines2 <- tran_lines %>% 
  filter(site_id %in% sites2)

#Transect start and end points and distance markers
ts_pts1 <- ts_pts %>% 
  filter(site_id %in% sites1)
ts_pts2 <- ts_pts %>% 
  filter(site_id %in% sites2)
te_pts1 <- te_pts %>% 
  filter(site_id %in% sites1)
te_pts2 <- te_pts %>% 
  filter(site_id %in% sites2)
dist_mark1 <- dist_mark %>% 
  filter(site_id %in% sites1)
dist_mark2 <- dist_mark %>% 
  filter(site_id %in% sites2)

#Seed load values
trees_sl1 <- trees_sl %>% 
  filter(site_id %in% sites1)
trees_sl2 <- trees_sl %>% 
  filter(site_id %in% sites2)

#Trees
trees_rot1 <- trees_rot %>% 
  filter(site_id %in% sites1)
trees_rot2 <- trees_rot %>% 
  filter(site_id %in% sites2)
```

##Thesis set 1
Plot first set
```{r}
p <- tmap_mode("plot") +
  #Define map extent using site footprints
  tm_shape(adjusted_fp_rot1, is.master = TRUE,
           unit = "m") +
    tm_facets(by = "site_id", free.scales = TRUE, 
              free.coords = TRUE, ncol = 2) +
    tm_polygons(col="white", border.col = NULL) +
  #Edge lines
  tm_shape(edge_lines1) + 
    tm_lines(col="blue", lty=2) +
  #Transect footprints and distance markers
  tm_shape(tran_lines1) + 
    tm_polygons(alpha=0) + 
  #Transect start and end points
  tm_shape(ts_pts1) + 
    tm_symbols(col="blue", shape=3, size=0.1) + 
  tm_shape(te_pts1) + 
    tm_symbols(col="blue", shape=3, size=0.1) +
  tm_shape(dist_mark1) + 
    tm_lines(lty = 3, alpha = 0.7) +
  #Seed load values
  tm_shape(trees_sl1) +
    tm_symbols(col = "sl_i_tot_rm", 
               palette = "Greys", n = 5, contrast = c(.1, 1),
               scale = 1,
               alpha = 0.6,
               style = "cont",
               border.lwd = NA,
               legend.col.is.portrait = T,
               title.col = "Seed load") +
  #Tree points
  tm_shape(trees_rot1) + 
    tm_symbols(col ="dmr_f_graph",
               palette = color_dmr,
               shape =  "sim_tree_g",
               shapes =  shape_sim, 
               size = "dbh",
               scale = 2,
               title.col = "HDM infection",
               legend.col.is.portrait = T,
               title.size = "DBH (cm)",
               legend.size.is.portrait = T,
               title.shape = "Tree type",
               legend.shape.is.portrait = T,
               shapes.legend.fill = "black",
               sizes.legend = c(10, 50, 100, 150)) +
  #Scale bar
  tm_scale_bar(position = c("RIGHT", "BOTTOM"), 
               lwd = 0.5,
               breaks = c(0, 25, 50),
               text.size = 1) +
  #Formatting
  tm_layout(inner.margins = 0.01, 
            panel.label.size = 1.3,
            panel.label.bg.color = "white",
            legend.outside = T, 
            legend.outside.position = "right",
            legend.position=c(0.1, 0.1), 
            legend.stack="vertical",
            legend.title.size = 1.2,
            legend.text.size = 1,
            legend.bg.color = "white",
            legend.bg.alpha = 1,
            legend.frame = T,
            fontfamily = "Times New Roman")

#Save the plot
# tmap_save(p, filename = "./figures/trees_sl_set1.svg", 
#           width = 6.5, height = 7.5, units = "in")
```

##Thesis set 2
Plot the second set
```{r}
p <- tmap_mode("plot") +
  #Define map extent using site footprints
  tm_shape(adjusted_fp_rot2, is.master = TRUE,
           unit = "m") +
    tm_facets(by = "site_id", free.scales = TRUE, 
              free.coords = TRUE, ncol = 2) +
    tm_polygons(col="white", border.col = NULL) +
  #Edge lines
  tm_shape(edge_lines2) + 
    tm_lines(col="blue", lty=2) +
  #Transect footprints and distance markers
  tm_shape(tran_lines2) + 
    tm_polygons(alpha=0) + 
  #Transect start and end points
  tm_shape(ts_pts2) + 
    tm_symbols(col="blue", shape=3, size=0.1) + 
  tm_shape(te_pts2) + 
    tm_symbols(col="blue", shape=3, size=0.1) +
  tm_shape(dist_mark2) + 
    tm_lines(lty = 3, alpha = 0.7) +
  #Seed load values
  tm_shape(trees_sl2) +
    tm_symbols(col = "sl_i_tot_rm", 
               palette = "Greys", n = 5, contrast = c(.1, 1),
               scale = 1,
               alpha = 0.6,
               style = "cont",
               border.lwd = NA,
               legend.col.is.portrait = T,
               title.col = "Seed load") +
  #Tree points
  tm_shape(trees_rot2) + 
    tm_symbols(col ="dmr_f_graph",
               palette = color_dmr,
               shape =  "sim_tree_g",
               shapes =  shape_sim, 
               size = "dbh",
               scale = 2,
               title.col = "HDM infection",
               legend.col.is.portrait = T,
               title.size = "DBH (cm)",
               legend.size.is.portrait = T,
               title.shape = "Tree type",
               legend.shape.is.portrait = T,
               shapes.legend.fill = "black",
               sizes.legend = c(10, 50, 100, 150)) +
  #Scale bar
  tm_scale_bar(position = c("RIGHT", "BOTTOM"), 
               lwd = 0.5,
               breaks = c(0, 25, 50),
               text.size = 1) +
  #Formatting
  tm_layout(inner.margins = 0.01, 
            panel.label.size = 1.3,
            panel.label.bg.color = "white",
            legend.outside = T, 
            legend.outside.position = "right",
            legend.position=c(0.1, 0.1), 
            legend.stack="vertical",
            legend.title.size = 1.2,
            legend.text.size = 1,
            legend.bg.color = "white",
            legend.bg.alpha = 1,
            legend.frame = T,
            fontfamily = "Times New Roman")

#Save the plot
# tmap_save(p, filename = "./figures/trees_sl_set2.svg", 
#           width = 6.5, height = 7.5, units = "in")
```

##Defence presentation plot
Plot three representative sites—a "high infection" (cr_3), "low infection" 
(ph_1) and average site (mi_1). 

Filter all plotting objects to these sites. 
```{r}
sites3 <- c("cr_3", "mi_1", "ph_1")

#Footprints
adjusted_fp_rot3 <- adjusted_fp_rot %>% 
  filter(site_id %in% sites3)

#Edge lines
edge_lines3 <- edge_lines %>% 
  filter(site_id %in% sites3)

#Transect footprints
tran_lines3 <- tran_lines %>% 
  filter(site_id %in% sites3)

#Transect start and end points and distance markers
ts_pts3 <- ts_pts %>% 
  filter(site_id %in% sites3)
te_pts3 <- te_pts %>% 
  filter(site_id %in% sites3)
dist_mark3 <- dist_mark %>% 
  filter(site_id %in% sites3)

#Seed load values
trees_sl3 <- trees_sl %>% 
  filter(site_id %in% sites3)

#Trees
trees_rot3 <- trees_rot %>% 
  filter(site_id %in% sites3)
```

Plot it
```{r}
#Rename site id by some text labels that are easier to interpret
adjusted_fp_rot3 <- adjusted_fp_rot3 %>% 
  mutate(site_id = case_match(site_id,
                              "cr_3" ~ "High infection",
                              "mi_1" ~ "Average infection",
                              "ph_1" ~ "Low infection")) %>% 
  mutate(site_id = factor(site_id, levels = c("High infection", 
                                              "Average infection",
                                              "Low infection")))

p <- tmap_mode("plot") +
  #Define map extent using site footprints
  tm_shape(adjusted_fp_rot3, is.master = TRUE,
           unit = "m") +
    tm_facets(by = "site_id", free.scales = TRUE, 
              free.coords = TRUE, ncol = 3) +
    tm_polygons(col="white", border.col = NULL) +
  #Edge lines
  tm_shape(edge_lines3) + 
    tm_lines(col="blue", lty=2) +
  #Transect footprints and distance markers
  tm_shape(tran_lines3) + 
    tm_polygons(alpha=0) + 
  #Transect start and end points
  tm_shape(ts_pts3) + 
    tm_symbols(col="blue", shape=3, size=0.1) + 
  tm_shape(te_pts3) + 
    tm_symbols(col="blue", shape=3, size=0.1) +
  tm_shape(dist_mark3) + 
    tm_lines(lty = 3, alpha = 0.7) +
  #Seed load values
  tm_shape(trees_sl3) +
    tm_symbols(col = "sl_i_tot_rm", 
               palette = "Greys", n = 5, contrast = c(.1, 1),
               scale = 1,
               alpha = 0.6,
               style = "cont",
               border.lwd = NA,
               legend.col.is.portrait = T,
               title.col = "Seed load") +
  #Tree points
  tm_shape(trees_rot3) + 
    tm_symbols(col ="dmr_f_graph",
               palette = color_dmr,
               shape =  "sim_tree_g",
               shapes =  shape_sim, 
               size = "dbh",
               scale = 2,
               title.col = "HDM infection",
               legend.col.is.portrait = T,
               title.size = "DBH (cm)",
               legend.size.is.portrait = T,
               title.shape = "Tree type",
               legend.shape.is.portrait = T,
               shapes.legend.fill = "black",
               sizes.legend = c(10, 50, 100)) +
  #Scale bar
  tm_scale_bar(position = c("RIGHT", "BOTTOM"), 
               lwd = 0.5,
               breaks = c(0, 25, 50),
               text.size = 1) +
  #Formatting
  tm_layout(inner.margins = 0.01, 
            panel.label.size = 1.3,
            panel.label.bg.color = "white",
            legend.outside = T, 
            legend.outside.position = "bottom",
            legend.position=c(0.35, 0.05), 
            legend.stack="horizontal",
            legend.title.size = 1,
            legend.text.size = 0.8,
            legend.bg.color = "white",
            legend.bg.alpha = 1,
            legend.frame = T,
            legend.outside.size = 0.3,
            fontfamily = "Times New Roman")

#Save the plot
# tmap_save(p, filename = "./figures/trees_sl_defence.svg",
#           width = 30, height = 15, units = "cm")
```

##Plot for WIFDWC 2024
```{r}
#Plot three representative sites: cr_3, mk_2, ph_1
#Filter all objects to these three sites
rep_sites <- c("cr_3", "mk_2", "ph_1")
adjusted_fp_rot <- adjusted_fp_rot %>% filter(site_id %in% rep_sites)
edge_lines <- edge_lines %>% filter(site_id %in% rep_sites)
tran_lines <- tran_lines %>% filter(site_id %in% rep_sites)
tran_pts <- tran_pts %>% filter(site_id %in% rep_sites)
dist_mark <- dist_mark %>% filter(site_id %in% rep_sites)
trees_rot <- trees_rot %>% filter(site_id %in% rep_sites)

pdf(here("./figures/trees_wifdwc.pdf"), width = 11, height = 4)
tmap_mode("plot")
#define map extent using site footprints
tm_shape(adjusted_fp_rot, is.master = TRUE) + #define map extent using site footprints
  tm_facets(by = "id_bec_ag_f", free.scales = FALSE, free.coords = TRUE, ncol = 3) +
  tm_polygons(col="white", border.col = NULL) +
#add edge lines
tm_shape(edge_lines) + 
  tm_lines(col="blue", lty=2) + 
#add transect pologyons, start/end points and labels
tm_shape(tran_lines) + 
  tm_polygons(alpha=0.25) + 
  tm_text("tr_leng", ymod = 1, scale = 0.8) + 
tm_shape(tran_pts) + 
  tm_symbols(col="blue", shape=3, size=0.1) + 
tm_shape(dist_mark) + tm_lines(lty = 3, alpha = 0.7) +
#add tree points
  tm_shape(trees_rot) + 
  tm_symbols(size = "dbh", scale = 1.5, col="dmr_f_graph", palette = dmr_colour, 
             border.lwd = NA, title.size = "DBH (cm)", title.col = "Modified DMR", 
             legend.size.is.portrait = TRUE, legend.col.is.portrait = TRUE) + 
  tm_layout(legend.outside=TRUE, 
            legend.outside.position = "right", 
            legend.position=c(0.05, 0.05),
            legend.stack="vertical",
            legend.frame = TRUE,
            legend.outside.size = .3,
            legend.title.size = 1.5,
            legend.text.size = 1,
            panel.label.size = 1.5)
dev.off()
```

