---
title: "Stem Mapped Figures"
author: "Hanno Southam"
date: "`r Sys.Date()`"
output: html_document
---
Read in data and load packages. 
"trees.geojson" is the output of the convert_distaz_to_points.R script. It contains all of the mature and regen trees as points and their associated data (e.g. DMR).  
"hdm_trimbpoints_2023.csv" are the trimble points from 2023 that have been cleaned up in GIS following the workflow in: /Users/hannosoutham/OneDrive - UBC (1)/Msc/GIS/ArcGIS_data_analysis_v1/hdm_arcgis_workflow.docx
"transect data_c.csv" is the cleaned transect data. Its used here to add the transect distances to the plot. 
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(RColorBrewer)

trees <- read_sf("./exports/trees.geojson")
trimb <- read_csv("./exports/trimb_processed.csv")
tran <- read_csv('./data/cleaned/transect data_c.csv')
?read_sf
```

Format the trees dataset. 
```{r}
class(trees)
summary(trees)

#check CRS, should be BC Albers NAD83 
st_crs(trees) #check CRS, should be BC Albers NAD83 = 3005

#Sort by site_id
trees <- trees %>% arrange(site_id)

# Convert to factor: spp, status, hdm_pa, b_lc, broom_pa, broom_stem, crown class, crown_cond, outside_10
trees <- trees %>% mutate(across(c(site_id, spp, status, hdm_pa, b_lc, broom_pa, broom_pos, broom_stem, crown_class, crown_cond, outside_10), ~ as.factor(.)))

#Convert DMR ratings to numerics.  
trees <- trees %>%
  mutate(across(c(dmr_l, dmr_m, dmr_u), ~ case_when(. == "-" ~ as.numeric(NA),
    TRUE ~ as.numeric(.))))
str(trees) #check

#Add a tree level dmr = dmr_l + dmr_m + dmr_u
trees <- trees %>% mutate(dmr = (dmr_l + dmr_m + dmr_u))

#Make another tree level DMR factor variable for plotting
trees <- trees %>% mutate(dmr_f = case_when(is.na(dmr) ~ "not Hw", dmr == 0 ~ "0, uninfected", dmr %in% c(1, 2) ~ "1-2, lightly infected", dmr %in% c(3, 4) ~ "3-4, moderately infected", dmr %in% c(5, 6) ~ "5-6, severely infected", TRUE ~ "case_when error")) %>% mutate(dmr_f=as.factor(dmr_f))
?case_when
```

Format the trimb dataset
```{r}
class(trimb)
summary(trimb)

#Convert to sf object
trimb <- st_as_sf(trimb, coords = c("plot_x_utm", "plot_y_utm"), crs=3005)
class(trimb)

#Check CRS is BC Albers NAD83 = 3005
st_crs(trimb)

#Sort by site_id
trimb <- trimb %>% arrange(site_id)

#Convert to pt_type and site_id to factor
trimb <- trimb %>% mutate(across(c(site_id, pt_type), ~ as.factor(.)))
str(trimb$pt_type)
str(trimb$site_id)

#Convert pt_id to a numeric
trimb <- trimb %>% mutate(pt_id = as.numeric(pt_id))
```

Rotate the points at each site (in both trees and trimb) so they are all oriented with the mature component running vertically and the transects running horizontally. Code adapted with help from Chat GPT from: https://stackoverflow.com/questions/31873151/how-rotate-map-in-r
```{r}
#Calculate vector with how many degrees each site needs to be rotated. Want each transect to be az = 90deg. 
degrees <- trees %>% 
  mutate(rot_deg = case_when(tr_az<90 ~ (-1)*(tr_az-90), #define degree adjustment
                             tr_az>90 ~ (-1)*(tr_az-90))) %>% 
  filter(str_detect(tree_id_new, "r")) %>% #filter to just regen trees
  group_by(site_id) %>% 
  distinct(rot_deg) %>% # get unique values for each site
  pull(rot_deg) #make vector

#Define a list of centre coordinate corresponding to each site. Use the centroid of all the stem mapped trees.
center_coords <- trees %>% group_by(site_id) %>% 
  summarise(geometry = st_centroid(st_combine(geometry))) %>% st_geometry()

#Split datasets, one for each site. This results in list of sf objects, one for each site. For the point datasets (trees and trimb) all the points at a site have been combined to a single feature (corresponding to a sf class change of POINT to MULTIPOINT). 
pts_trees <- trees %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()
pts_trimb <- trimb %>% group_by(site_id) %>% 
  summarise(geometry=st_combine(geometry)) %>% st_geometry()

#Define a function to rotate a spatial feature. x = an sf object, center_coords = the centre coordinates to anchor each site, degrees=the number of degrees to rotate the object
rotate_sf <- function(x, center_coords, degrees) {
  radians <- degrees * pi/180 #convert to radians
  transform_matrix <- matrix(c(cos(radians), sin(radians), #the transformation matrix
                                 -sin(radians), cos(radians)), 2, 2)
  rot_site <- ((x - center_coords) * transform_matrix) + center_coords #the actual math
  return(rot_site) #output
}

#Map function applies another function to each item in a vector successively. Here, it rotates each site level MULTIPOINT sf object by the number of degrees in the "degrees" vector. 
trees_rot <- Map(rotate_sf, pts_trees, center_coords, degrees)

trimb_rot <- Map(rotate_sf, pts_trimb, center_coords, degrees)

#This output is a new list of sf multipoint objects. Combine these back together into a dataframe of x, y coordinates. 
trees_rot <- do.call(rbind, trees_rot) #do.call applies function to each item in a list
trees_rot <- as.data.frame(trees_rot) %>% rename(X=V1, Y=V2)

trimb_rot <- do.call(rbind, trimb_rot) #do.call applies function to each item in a list
trimb_rot <- as.data.frame(trimb_rot) %>% rename(X=V1, Y=V2)

#Then associate back to the attribute data associated with each point
trees_att <- trees %>% st_drop_geometry()
trees_rot <- trees_att %>% mutate(X=trees_rot$X, Y=trees_rot$Y) 
trees_rot <- st_as_sf(trees_rot, coords = c("X", "Y"), crs=3005)

trimb_att <- trimb %>% st_drop_geometry()
trimb_rot <- trimb_att %>% mutate(X=trimb_rot$X, Y=trimb_rot$Y) 
trimb_rot <- st_as_sf(trimb_rot, coords = c("X", "Y"), crs=3005)
```

Make the stem mapped plot. 
```{r}
#Create the footprints of the each site:
tran_pts <- trimb_rot %>% #filter trimb_points to just transect starts and ends
  filter(pt_type %in% c("r adj tran start", "tran end"))
rot_points <- rbind(select(trees_rot, site_id), select(tran_pts, site_id)) #combine trees_rot and trimb_rot datasets
rot_points <- rot_points %>% group_by(site_id) %>% #split rotated points into separate datasets for each site
  summarise(geometry=st_combine(geometry)) %>% st_geometry()
fp_rot <- Map(st_bbox, rot_points) #create bounding box around each of these sets of points
fp_rot <- Map(st_as_sfc, fp_rot) #turn bbox object into sfc object
fp_rot <- Map(st_as_sf, fp_rot) #trun sfc object into sf object
fp_rot <- do.call(rbind, fp_rot) #combine sf objects to single object
sites <- trees %>% st_drop_geometry() %>% distinct(site_id) #get column of site_ids
fp_rot <- fp_rot %>% mutate(site_id=sites$site_id) #add comlumn of site_ids to sf object
st_crs(fp_rot) = 3005 #define crs

#Create the transect lines for the site
tran_pts <- trimb_rot %>% #filter to just transect starts and ends
  filter(pt_type %in% c("r adj tran start", "tran end"))
tran_line_att <- tran_pts %>% st_drop_geometry() #save attributes
tran_lines <- tran_pts %>% group_by(pt_id) %>% # draw lines between start and end. Creates two lines one from start to end and one from end to start.
  summarize() %>% st_cast("LINESTRING")
tran_lines <- st_buffer(tran_lines, dist = 2.5, endCapStyle = 'FLAT') # buffer
tran_lines <- left_join(tran_lines, tran_line_att, by="pt_id")
tran_lines <- tran_lines %>% filter(pt_type=="tran end")
tran_lines <- left_join(tran_lines, select(tran, transect_id, tr_leng), by=c("pt_id"="transect_id"))
?st_cast

#Create some edge lines: Can't figure out how to plot in right line
levels(trimb_rot$pt_type)
edge_lines <- trimb_rot %>% #filter to points that define edge ("edge" and "r adj tran start")
  filter(pt_type %in% c("edge", "r adj tran start"))
edge_line_coords <- st_coordinates(edge_lines) %>%  #extract coordinates
  as.data.frame() %>% mutate(diff=(X-Y))
edge_lines <- edge_lines %>% mutate(diff=edge_line_coords$diff)
edge_lines <- edge_lines %>% group_by(site_id) %>% arrange(diff) %>% # draw lines between start and end
  summarise(do_union = FALSE) %>% st_cast("LINESTRING")

#Define colour pallette
dmr_colour <- c('#2171b5','#fed98e','#fe9929','#cc4c02', '#cccccc')

# Plot it all together
png("./figures/trees_27Mar2024.png", width=3300, height=2550, units="px", res = 300)
tmap_mode("plot")
tm_shape(fp_rot, is.master = TRUE) + #define map extent using site footprints
  tm_facets(by="site_id", free.scales = FALSE, free.coords = TRUE, ncol = 4) +
  tm_polygons(col="white", border.col = NULL) +
tm_shape(edge_lines) + tm_lines(col="blue", lty=2) + #add edge lines
tm_shape(tran_lines) + tm_polygons(alpha=0.75) + 
  tm_text("tr_leng", ymod = 1.7) + #add transects
tm_shape(tran_pts) + tm_symbols(col="blue", shape=3, size=0.1) + #transect start points
tm_shape(trees_rot) + tm_symbols(size = "dbh", col="dmr_f", palette = dmr_colour) + #add tree points
tm_legend(legend.outside=TRUE, legend.outside.position="bottom", 
          legend.position = c(0.4, 0.1), legend.stack="horizontal")
dev.off()
```

